---
title: "Birth decline in Switzerland: "
format: html
editor: visual
execute:
  warning: false
  message: false
  echo: false
---

# Data

```{r}
#| include: false
source("R/000_setup.R")

load("results/res.RData")
```


# Models

## Model 1

### Non-parametric age-distribution of births

By age and by year from 2000 to 2024: $25 \times 36 = 900$ rows.<br> Neg-binomial distribution of births with an offset on the population.<br> Effect of years and of age on birth rates modelled with Gaussian processes (GPs).<br> Limitations: For a given age, cannot disentangle the variation in birth rate due to increase in age of delivery or genuine change in number of births.

#### Same effect of year for any year

```{r}
#| fig-width: 10
#| fig-height: 6

cowplot::plot_grid(
  #GP age
  res1$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res1$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```

#### Age-specific effect of year

```{r}
#| fig-width: 10
#| fig-height: 6
cowplot::plot_grid(
  #GP age
  res2$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res2$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```

### Parametric age-distribution of births

Parametric function to model the probability of giving birth with respect to age (Gaussian curve).<br> Three parameters govern the Gaussian curve: 1) peak age, 2) peak height and 3) width (i.e., concentration around peak).

#### Time-varying peak age

```{r}
#parametric function
res3$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res3$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res3$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

### Model 2 bis

```{r}
#parametric function
res4$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res4$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

### Semi-parametric age-distribution of births

Semi-parametric function to model the probability of giving birth with respect to age (random walk).<br>

#### Time-varying peak age

```{r}
#parametric function
res6$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res6$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res6$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

## Model 2

By age, year and months: $12\times 25 \times 36  = 10,800$ rows.<br> Seasonal variation captured by a periodic GP.

```{r}
#parametric function
res5$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024),month==1) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))

```

```{r}
#| fig-width: 8
#| fig-height: 7
#GP
cowplot::plot_grid(
res5$gp_year_df %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),#facet_grid(group_id ~.)
res5$gp_month_df %>%
  filter(year<=2001) %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),ncol=1)
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

## Model 3

By age, year, months and cantons: $12\times 25 \times 36 \times 26 = 280,800$ rows.
This was abandoned because we want comparable estimates over canton, i.e., where the expected (mean) birth does not depend on canton.



# Main points (old)

**Objective**

-   2 parts:
    -   Understand long-term evolution of birth in Switzerland (peak age) and regional differences: expectation
    -   Identify and explain (i.e., adjust for some confounding) birth excess/deficit.

**Stratification**

-   Stratifications increase number of likelihood estimates, which increases computation time
-   Even if level of stratification in the model is coarse (e.g., national level), we will still be able to perform some post-processing where we compute excess/deficit in birth by finer levels (e.g., by district, municipality, SEP, etc)
-   The models, through their different levels of stratification, serve different purposes:
    -   National-level expectation:
        -   Inspect regional deviations
    -   Regional-level expectation:
        -   Explore finer geographical effect (municipality, district)
        -   Explore effect of SEP after accounting for potential confounding by region.

**Modelling age distribution**

-   Why explicitly modelling mother age distribution:
    -   to quantify age shift
    -   to disentangle effects of age shift and of genuine fertility change on birth rate
    -   to adjust for confounding by population age to isolate the effect of region on birth rate (e.g., a region with more 30-year-old women is expected to have higher fertility rates)
-   Bias in birth expectation by age due to the use to a parametric function lacking flexibility: try more flexible functions
-   Some effect might be overfitted (e.g., time-varying time peak height).

**To explore**

-   Effect of nationality could be interesting to explore.

**Questions** - How to filter: live birth, permanent population

# Issues

-   Using a Gaussian curve introduces biases at some ages and does not capture the true mean properly.
-   A warm-up of 200 iterations is not sufficient and can lead to divergences.
-   Some divergences (around 1%) are still observed when `adapt_delta = 0.8`.
-   Using a negative-binomial parameterization with `var = mu * (1 + phi)` leads to many divergences (more than 80%).
-   One row with a Swiss mother is missing `mun_id` (currently removed); analysis is restricted to `mother_permanent == 1` and `live_birth == 1`.

# To do

-   SEP: can only for 2010-2024 because we need population by municipality to assign SEP.
-   Launch in cluster, parallelization possible

