---
title: "Birth decline in Switzerland: "
format: html
editor: visual
execute:
  warning: false
  message: false
  echo: false
---

# Data

```{r}
#| include: false
source("R/000_setup.R")

#load data
#birth data, 1987_2024
birth_1987_2024 = load_birth_data()
birth_df = birth_1987_2024 %>% 
  filter(live_birth==1,mother_permanent==1) %>% 
  dplyr::select(year,month,mother_age,birth_loc,birth_state,mother_municipality,mother_citizenship,parity) %>% 
  dplyr::mutate(mother_birth_year = year-mother_age)
```


## Population data

Stratification level: age, year (1987-2024), canton

- Restrict to women between 15 and 50.
- Linear interpolation within months
- Finer level of stratification (e.g., municipality) for 2010-2024

## Birth data

Stratification level: individual-based

- increase in high age and decrease in low age over time
- probability of giving birth by age over time
- municipality

```{r}

```

```{r}

```

### Number of births

#### Over time

```{r}
load("results/res.RData")
#Total number of births---------------------------------------------------------
#by year: decrease from 2016, exception in 2021
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n)) + geom_point()+geom_line()+
  expand_limits(y = 0)

```

#### By month

```{r}
birth_df %>% 
  filter(year %in% c(1990,2000,2010,2020,2024)) %>% 
  group_by(month,year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=month,y=n,col=factor(year))) + geom_point()+geom_line()
```

#### By citizenship

```{r}
#number of births by mother citizenship
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship==8100)) %>%
  group_by(year,mother_citizenship) %>% dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n,col=factor(mother_citizenship)))+geom_point()+geom_line()+
  expand_limits(y = 0)
```

### Mother age

#### Mean mother age over time

```{r}
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```


#### Mother age distribution by calendar year

```{r}
birth_df |>
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Year")
```


#### Mean mother age by citizenship

```{r}
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  group_by(year, mother_citizenship) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean, col = factor(mother_citizenship))) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```

#### Mother age distribution by calendar year and citizenship

```{r}
#| fig-width: 8
#| fig-height: 10
birth_df |>
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_citizenship))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Swiss") +
  facet_grid(year ~ .)+
  theme(legend.position = "bottom")
```

#### Mother age by mother birth year (right-censoring adjusted)

```{r}
years <- c(1970, 1975, 1980, 1985, 1990, 1995)

birth_df %>% 
  filter(mother_birth_year %in% years,
         mother_age > (1987 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year")
```

#### Mother age (first child only, density)

```{r}
years <- c(1982, 1983, 1988, 1990)

birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(parity == 1,
         mother_birth_year %in% years,
         mother_age > (2004 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year") +
  facet_grid(mother_citizenship ~ .)
```

### Parity

#### Parity distribution over selected years

```{r}
birth_df %>% 
  filter(year %in% c(2005, 2010, 2020, 2024),
         parity <= 5) %>% 
  ggplot(aes(x = parity,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Parity", y = "Count", color = "Year")
```

#### Proportion of parity 2–4 over time

```{r}
birth_df %>% 
  filter(year >= 2005) %>% 
  group_by(year, parity) %>% 
  dplyr::summarise(n = n(), .groups = "drop_last") %>% 
  dplyr::mutate(p = n / sum(n)) %>% 
  ungroup() %>% 
  filter(parity >= 2, parity <= 4) %>% 
  ggplot(aes(x = year,
             y = p,
             col = factor(parity))) +
  geom_line() +
  geom_point() +
  expand_limits(y = 0)
```

# Models

## Model 1

### Non-parametric age-distribution of births

By age and by year from 2000 to 2024: $25 \times 36 = 900$ rows.<br>
Neg-binomial distribution of births with an offset on the population.<br>
Effect of years and of age on birth rates modelled with Gaussian processes (GPs).<br>
Limitations: For a given age, cannot disentangle the variation in birth rate due to increase in age of delivery or genuine change in number of births.

#### Same effect of year for any year

```{r}
#| fig-width: 10
#| fig-height: 6
cowplot::plot_grid(
  #GP age
  res1$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res1$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```

#### Age-specific effect of year

```{r}
#| fig-width: 10
#| fig-height: 6
cowplot::plot_grid(
  #GP age
  res2$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res2$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```


### Parametric age-distribution of births

Parametric function to model the probability of giving birth with respect to age (Gaussian curve).<br>
Three parameters govern the Gaussian curve: 1) peak age, 2) peak height and 3) width (i.e., concentration around peak).

#### Time-varying peak age

```{r}
#parametric function
res3$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res3$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res3$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

### Model 2 bis

```{r}
#parametric function
res4$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res4$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

### Semi-parametric age-distribution of births

Semi-parametric function to model the probability of giving birth with respect to age (random walk).<br>

#### Time-varying peak age

```{r}
#parametric function
res6$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res6$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res6$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

## Model 2

By age, year and months: $12\times 25 \times 36  = 10,800$ rows.<br>
Seasonal variation captured by a periodic GP.

```{r}
#parametric function
res5$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024),month==1) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))

```


```{r}
#| fig-width: 8
#| fig-height: 7
#GP
cowplot::plot_grid(
res5$gp_year_df %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),#facet_grid(group_id ~.)
res5$gp_month_df %>%
  filter(year<=2001) %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),ncol=1)
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```


```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

## Model 3

By age, year, months and cantons: $12\times 25 \times 36 \times 26 = 280,800$ rows.

## Model 4

By age, year, months, cantons and SEP: $12\times 25 \times 36 \times 26 \times 3 = 842,400$ rows.
But only for 2010-2024.

# Main points

**Objective**

- 2 parts:
  - Understand long-term evolution of birth in Switzerland (peak age) and regional differences: expectation
  - Identify and explain (i.e., adjust for some confounding) birth excess/deficit.

**Stratification**

- Stratifications increase number of likelihood estimates, which increases computation time
- Even if level of stratification in the model is coarse (e.g., national level), we will still be able to perform some post-processing where we compute excess/deficit in birth by finer levels (e.g., by district, municipality, SEP, etc)
- The models, through their different levels of stratification, serve different purposes:
  - National-level expectation:
    - Inspect regional deviations
  - Regional-level expectation:
    - Explore finer geographical effect (municipality, district)
    - Explore effect of SEP after accounting for potential confounding by region.

**Modelling age distribution**

- Why explicitly modelling mother age distribution:
  - to quantify age shift
  - to disentangle effects of age shift and of genuine fertility change on birth rate
  - to adjust for confounding by population age to isolate the effect of region on birth rate (e.g., a region with more 30-year-old women is expected to have higher fertility rates)
- Bias in birth expectation by age due to the use to a parametric function lacking flexibility: try more flexible functions
- Some effect might be overfitted (e.g., time-varying time peak height).

**To explore**

- Effect of nationality could be interesting to explore.

**Questions**
- How to filter: live birth, permanent population

# Issues

- Using a Gaussian curve introduces biases at some ages and does not capture the true mean properly.
- A warm-up of 200 iterations is not sufficient and can lead to divergences.
- Some divergences (around 1%) are still observed when `adapt_delta = 0.8`.
- Using a negative-binomial parameterization with `var = mu * (1 + phi)` leads to many divergences (more than 80%).
- One row with a Swiss mother is missing `mun_id` (currently removed); analysis is restricted to `mother_permanent == 1` and `live_birth == 1`.

# To do

- SEP: can only for 2010-2024 because we need population by municipality to assign SEP.
- Launch in cluster, parallelization possible
