---
title: "Birth decline in Switzerland: "
format: html
editor: visual
execute:
  warning: false
  message: false
  echo: false
---

# Data

```{r}
#| include: false
source("R/000_setup.R")

#load data
#birth data, 1987_2024
birth_1987_2024 = load_birth_data()
birth_df = birth_1987_2024 %>% 
  filter(live_birth==1,mother_permanent==1) %>% 
  dplyr::select(year,month,mother_age,birth_loc,birth_state,mother_municipality,mother_citizenship,parity) %>% 
  dplyr::mutate(mother_birth_year = year-mother_age)
```

## Population data

Stratification level: age, year (1987-2024), canton

-   Restrict to women between 15 and 50.
-   Linear interpolation within months
-   Finer level of stratification (e.g., municipality) for 2010-2024

## Birth data

Stratification level: individual-based

-   increase in high age and decrease in low age over time
-   probability of giving birth by age over time
-   municipality

```{r}

```

```{r}

```

### Number of births

#### Over time

```{r}
load("results/res.RData")
#Total number of births---------------------------------------------------------
#by year: decrease from 2016, exception in 2021
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n)) + geom_point()+geom_line()+
  expand_limits(y = 0)

```

#### By month

```{r}
birth_df %>% 
  filter(year %in% c(1990,2000,2010,2020,2024)) %>% 
  group_by(month,year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=month,y=n,col=factor(year))) + geom_point()+geom_line()
```

#### By citizenship

```{r}
#number of births by mother citizenship
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship==8100)) %>%
  group_by(year,mother_citizenship) %>% dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n,col=factor(mother_citizenship)))+geom_point()+geom_line()+
  expand_limits(y = 0)
```

### Mother age

#### Mean mother age over time

```{r}
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```

#### Mother age distribution by calendar year

```{r}
birth_df |>
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Year")
```

#### Mean mother age by citizenship

```{r}
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  group_by(year, mother_citizenship) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean, col = factor(mother_citizenship))) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```

#### Mother age distribution by calendar year and citizenship

```{r}
#| fig-width: 8
#| fig-height: 10
birth_df |>
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_citizenship))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Swiss") +
  facet_grid(year ~ .)+
  theme(legend.position = "bottom")
```

#### Mother age by mother birth year (right-censoring adjusted)

```{r}
years <- c(1970, 1975, 1980, 1985, 1990, 1995)

birth_df %>% 
  filter(mother_birth_year %in% years,
         mother_age > (1987 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year")
```

#### Mother age (first child only, density)

```{r}
years <- c(1982, 1983, 1988, 1990)

birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(parity == 1,
         mother_birth_year %in% years,
         mother_age > (2004 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year") +
  facet_grid(mother_citizenship ~ .)
```

### Parity

#### Parity distribution over selected years

```{r}
birth_df %>% 
  filter(year %in% c(2005, 2010, 2020, 2024),
         parity <= 5) %>% 
  ggplot(aes(x = parity,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Parity", y = "Count", color = "Year")
```

#### Proportion of parity 2–4 over time

```{r}
birth_df %>% 
  filter(year >= 2005) %>% 
  group_by(year, parity) %>% 
  dplyr::summarise(n = n(), .groups = "drop_last") %>% 
  dplyr::mutate(p = n / sum(n)) %>% 
  ungroup() %>% 
  filter(parity >= 2, parity <= 4) %>% 
  ggplot(aes(x = year,
             y = p,
             col = factor(parity))) +
  geom_line() +
  geom_point() +
  expand_limits(y = 0)
```

# Models

## Model 1

### Non-parametric age-distribution of births

By age and by year from 2000 to 2024: $25 \times 36 = 900$ rows.<br> Neg-binomial distribution of births with an offset on the population.<br> Effect of years and of age on birth rates modelled with Gaussian processes (GPs).<br> Limitations: For a given age, cannot disentangle the variation in birth rate due to increase in age of delivery or genuine change in number of births.

#### Same effect of year for any year

```{r}
#| fig-width: 10
#| fig-height: 6
cowplot::plot_grid(
  #GP age
  res1$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res1$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```

#### Age-specific effect of year

```{r}
#| fig-width: 10
#| fig-height: 6
cowplot::plot_grid(
  #GP age
  res2$gp_age_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(fill="blue",alpha=0.2)+
    geom_line(col="blue"),
  #GP year
  res2$gp_year_df %>% 
    ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(group_year_id)),alpha=0.2)+
    geom_line(aes(col=factor(group_year_id)))+
    theme(legend.position = "bottom"))
```

### Parametric age-distribution of births

Parametric function to model the probability of giving birth with respect to age (Gaussian curve).<br> Three parameters govern the Gaussian curve: 1) peak age, 2) peak height and 3) width (i.e., concentration around peak).

#### Time-varying peak age

```{r}
#parametric function
res3$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res3$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res3$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res3$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

### Model 2 bis

```{r}
#parametric function
res4$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res4$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

### Semi-parametric age-distribution of births

Semi-parametric function to model the probability of giving birth with respect to age (random walk).<br>

#### Time-varying peak age

```{r}
#parametric function
res6$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024)) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))
```

```{r}
#GP
res6$gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  facet_grid(group_id ~.)
```

#### Prediction intervals

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12

res6$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=year))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=2)+
      facet_grid(mother_age~.,scale="free_y")
```

#### Bias

```{r}
#| fig-width: 8
#| fig-height: 6

res6$age_bias_df %>% 
      ggplot(aes(x=mother_age))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      facet_grid(indicator~.,scales="free")
```

## Model 2

By age, year and months: $12\times 25 \times 36  = 10,800$ rows.<br> Seasonal variation captured by a periodic GP.

```{r}
#parametric function
res5$birth_prob_by_age_df %>% 
  filter(year %in% c(2000,2010,2021,2024),month==1) %>% 
  ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
  geom_line(aes(col=factor(year)))

```

```{r}
#| fig-width: 8
#| fig-height: 7
#GP
cowplot::plot_grid(
res5$gp_year_df %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),#facet_grid(group_id ~.)
res5$gp_month_df %>%
  filter(year<=2001) %>% 
  ggplot(aes(x=date,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue"),ncol=1)
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(15,16,18,20,22,24,26)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(28:34)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

```{r}
#| fig-width: 8
#| fig-height: 12
res5$pred_df %>% 
      filter(mother_age %in% c(36,38,40,42,45,48,50)) %>% 
      ggplot(aes(x=date))+
      geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
      geom_line(aes(y=est),col="darkred")+
      geom_point(aes(y=n_birth),size=1.5,alpha=0.5)+
      facet_grid(mother_age~.,scale="free_y")
```

## Model 3

By age, year, months and cantons: $12\times 25 \times 36 \times 26 = 280,800$ rows.

## Model 4

By age, year, months, cantons and SEP: $12\times 25 \times 36 \times 26 \times 3 = 842,400$ rows. But only for 2010-2024.

# Main points

**Objective**

-   2 parts:
    -   Understand long-term evolution of birth in Switzerland (peak age) and regional differences: expectation
    -   Identify and explain (i.e., adjust for some confounding) birth excess/deficit.

**Stratification**

-   Stratifications increase number of likelihood estimates, which increases computation time
-   Even if level of stratification in the model is coarse (e.g., national level), we will still be able to perform some post-processing where we compute excess/deficit in birth by finer levels (e.g., by district, municipality, SEP, etc)
-   The models, through their different levels of stratification, serve different purposes:
    -   National-level expectation:
        -   Inspect regional deviations
    -   Regional-level expectation:
        -   Explore finer geographical effect (municipality, district)
        -   Explore effect of SEP after accounting for potential confounding by region.

**Modelling age distribution**

-   Why explicitly modelling mother age distribution:
    -   to quantify age shift
    -   to disentangle effects of age shift and of genuine fertility change on birth rate
    -   to adjust for confounding by population age to isolate the effect of region on birth rate (e.g., a region with more 30-year-old women is expected to have higher fertility rates)
-   Bias in birth expectation by age due to the use to a parametric function lacking flexibility: try more flexible functions
-   Some effect might be overfitted (e.g., time-varying time peak height).

**To explore**

-   Effect of nationality could be interesting to explore.

**Questions** - How to filter: live birth, permanent population

# Issues

-   Using a Gaussian curve introduces biases at some ages and does not capture the true mean properly.
-   A warm-up of 200 iterations is not sufficient and can lead to divergences.
-   Some divergences (around 1%) are still observed when `adapt_delta = 0.8`.
-   Using a negative-binomial parameterization with `var = mu * (1 + phi)` leads to many divergences (more than 80%).
-   One row with a Swiss mother is missing `mun_id` (currently removed); analysis is restricted to `mother_permanent == 1` and `live_birth == 1`.

# To do

-   SEP: can only for 2010-2024 because we need population by municipality to assign SEP.
-   Launch in cluster, parallelization possible







```{r}
#national, by year--------------------------------------------------------------
excess_birth_year_df = pred_n_birth_draw_df %>% 
  #sum over age and month
  group_by(year,draw) %>% 
  dplyr::summarise(n_pop = sum(n_pop),
                   n_birth = sum(n_birth),
                   n_pred = sum(n_pred),.groups="drop_last") %>% 
  dplyr::summarise(n_pop = n_pop[1],
                   n_birth = n_birth[1],
                   n_exp_mean = mean(n_pred),
                   n_exp_lwb = quantile(n_pred,probs=0.025),
                   n_exp_upb = quantile(n_pred,probs=0.9725),.groups="drop") %>% 
  dplyr::mutate(n_exc_mean = n_birth - n_exp_mean,
                n_exc_lwb = n_birth- n_exp_lwb,
                n_exc_upb = n_birth-n_exp_upb,
                rel_exc_mean = (n_birth - n_exp_mean)/n_exp_mean,
                rel_exc_lwb = (n_birth- n_exp_lwb)/n_exp_mean,
                rel_exc_upb = (n_birth-n_exp_upb)/n_exp_mean)

plot_df = excess_birth_year_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~.,scale="free_y")

#national, by year and month----------------------------------------------------
excess_birth_year_month_df = pred_n_birth_reg_draw_df %>% 
  #sum over age and month
  group_by(year,month,draw) %>% 
  dplyr::summarise(n_pop = sum(n_pop),
                   n_birth = sum(n_birth),
                   n_pred = sum(n_pred),.groups="drop_last") %>% 
  dplyr::summarise(n_pop = n_pop[1],
                   n_birth = n_birth[1],
                   n_exp_mean = mean(n_pred),
                   n_exp_lwb = quantile(n_pred,probs=0.025),
                   n_exp_upb = quantile(n_pred,probs=0.9725),.groups="drop") %>% 
  dplyr::mutate(n_exc_mean = n_birth - n_exp_mean,
                n_exc_lwb = n_birth- n_exp_lwb,
                n_exc_upb = n_birth-n_exp_upb,
                rel_exc_mean = (n_birth - n_exp_mean)/n_exp_mean,
                rel_exc_lwb = (n_birth- n_exp_lwb)/n_exp_mean,
                rel_exc_upb = (n_birth-n_exp_upb)/n_exp_mean) %>% 
  dplyr::mutate(date = as.Date(sprintf("%d-%02d-01", year, month)))

plot_df = excess_birth_year_month_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year>=2020) %>% 
  ggplot(aes(x = date, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>%  filter(year>=2020) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>%  filter(year>=2020), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~.,scale="free_y")

#by region----------------------------------------------------------------------
excess_birth_year_reg_df = pred_n_birth_reg_draw_df %>% 
  #sum over age and month
  group_by(dist_name,dist_id,year,draw) %>% 
  dplyr::summarise(n_pop = sum(n_pop),
                   n_birth = sum(n_birth),
                   n_pred = sum(n_pred),.groups="drop_last") %>% 
  dplyr::summarise(n_pop = n_pop[1],
                    n_birth = n_birth[1],
                   n_exp_mean = mean(n_pred),
                   n_exp_lwb = quantile(n_pred,probs=0.025),
                   n_exp_upb = quantile(n_pred,probs=0.9725),.groups="drop") %>% 
  dplyr::mutate(n_exc_mean = n_birth - n_exp_mean,
                n_exc_lwb = n_birth- n_exp_lwb,
                n_exc_upb = n_birth-n_exp_upb,
                rel_exc_mean = (n_birth - n_exp_mean)/n_exp_mean,
                rel_exc_lwb = (n_birth- n_exp_lwb)/n_exp_mean,
                rel_exc_upb = (n_birth-n_exp_upb)/n_exp_mean))

#load region location
regions_sf <- st_read("data/boundary_data/Boundaries_K4_District_20260101.shp") #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
regions_sf = regions_sf[,c(1,3,7)]
names(regions_sf)[c(1,2)] <- c("dist_name2", "dist_id")

excess_birth_year_reg_df %>% 
  filter(year %in% 2017:2024) %>% 
  dplyr::mutate(rel_exc_mean = n_exc_mean/n_exp_mean) %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = rel_exc_mean)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(year~.,ncol=2)

#other
excess_birth_year_df %>% 
  ggplot(aes(x = year, y = n_exp_mean, ymin = n_exp_lwb, ymax = n_exp_upb)) +
  geom_ribbon( alpha = 0.2) +geom_line()
excess_birth_year_df %>% 
  ggplot(aes(x = year, y = n_exp_mean/n_pop, ymin = n_exp_lwb/n_pop, ymax = n_exp_upb/n_pop)) +
  geom_ribbon( alpha = 0.2) +geom_line()

#district
df = excess_birth_year_reg_df %>% 
  filter(dist_id %in% sample(unique(excess_birth_year_reg_df$dist_id), 6)) %>% 
  left_join(excess_birth_year_df %>% dplyr::select(year,n_pop_nat=n_pop,n_exp_mean_nat=n_exp_mean),by="year")
df %>% 
  ggplot(aes(x = year, y = n_exp_mean, ymin = n_exp_lwb, ymax = n_exp_upb)) +
  geom_ribbon(aes(fill=dist_name), alpha = 0.2) +
  geom_line(aes(col=dist_name)) +
  facet_grid(dist_name~.,scale="free_y")
df %>% 
  ggplot(aes(x = year, y = n_exp_mean/n_pop, ymin = n_exp_lwb/n_pop, ymax = n_exp_upb/n_pop)) +
  geom_ribbon(aes(fill=dist_name), alpha = 0.2) +
  geom_line(aes(col=dist_name)) +
  geom_line(aes(y=n_exp_mean_nat/n_pop_nat),lty=2)+
  facet_grid(dist_name~.)
excess_birth_year_reg_df %>% 
  ggplot(aes(x = year, y = n_exp_mean/n_pop)) +
  geom_line(aes(col=dist_name)) +
  expand_limits(y=0)+theme(legend.position = "none")

```
