---
title: "Birth decline in Switzerland: "
format: html
editor: visual
---

## Data

### Birth data

Stratification level: individual-based

- increase in high age and decrease in low age over time
- probability of giving birth by age over time
- municipality

### Population data

Stratification level: age, year (1987-2024), canton

- Restrict to women between 15 and 50.
- Linear interpolation within months
- Finer level of stratification (e.g., municipality) for 2010-2024



## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).


# Issues

- Warm-up of 200 is not sufficient, can create divergences
- Some divergences (around 1\%) also observed when adapt_delta=0.8
- When we use negative-binomial with var=mu*(1+phi), it leads to many divergence (more than 80\%)
- There is one row with Swiss mother missing mun_id (currently removed)

# To do

- SEP: can only for 2010-2024 because we need population by municipality to assign SEP.
- Launch in cluster, parallelization possible
