---
title: "Birth decline in Switzerland"
format:
  html:
    theme: cosmo
    code-fold: true        # adds "Show/Hide Code" buttons
    code-summary: "Code"   # customize button text
    toc: true
    toc-location: left  # place TOC on the left
    number-sections: true  # number sections
    toc-depth: 3      # show subsections up to 3 levels
    self-contained: true
    toc-fold: true     # allow collapsing/expanding sections
execute:
  echo: false
  fig-width: 7
  fig-height: 5
  warning: false
  message: true
  comment: ""
    
editor: visual
---


```{r,include=FALSE}
source("../R/000_setup.R")

#define models
mod_name = "mod8"
save.date = "20260309"
save.date2 = "20260320"
seed_id = 1

effect_on_age_shift = ifelse(grepl("birthyear", mod_name), "birth_year", "cal_year")
  
filter_parity = dplyr::case_when(grepl("second", mod_name) ~ "second",
                                 grepl("first", mod_name)  ~ "first",
                                 TRUE ~ "all")
  
filter_ctz2 = dplyr::case_when(grepl("non-swiss", mod_name) ~ "non-swiss",
                               grepl("swiss", mod_name) ~ "swiss",
                               TRUE ~ "all")
include_ctz = TRUE #filter_ctz2!="swiss"
```

```{r}

#load results
#stan_diag_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_standiag.RDS"))
#par_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_par.RDS"))
#gamma_month_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gammamonth.RDS"))
#b_year_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_byear.RDS"))
#sigma_year_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_sigmayear.RDS"))
birth_prob_by_age_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_birthprob.RDS"))
birth_prob_by_age_swiss_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","swiss","_seedid",seed_id,"_birthprob.RDS"))
birth_prob_by_age_nonswiss_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","non-swiss","_seedid",seed_id,"_birthprob.RDS"))

#gp_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gp.RDS"))
gp_rel_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gp_rel.RDS"))
#age_bias_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_agebias.RDS"))
#pred_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_pred.RDS"))

#load excess
excess_birth_year_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_swiss_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","swiss","_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_nonswiss_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","non-swiss","_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_first_df = readRDS(paste0(code_root_path,"results/",save.date2,"_",mod_name,"_","first","_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_second_df = readRDS(paste0(code_root_path,"results/",save.date2,"_",mod_name,"_","second","_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
#excess_birth_year_age_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_df",".RDS"))
excess_birth_year_month_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_df",".RDS"))
#excess_birth_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_reg_df",".RDS"))
#excess_birth_year_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_reg_df",".RDS"))

#excess by citizenship
if(filter_ctz2=="all"){
  excess_birth_year_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_ctz_df",".RDS"))
  #excess_birth_year_age_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_ctz_df",".RDS"))
  #excess_birth_year_month_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_ctz_df",".RDS"))
}

#excess by municipality
#excess_birth_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_mun_df",".RDS"))
excess_birth_adj2_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_adj2_mun_df",".RDS"))
if(filter_parity!="all"){
  excess_birth_adj2_mun_childless_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_childless","_","seedid",seed_id,"_","excess_birth_adj2_mun_df",".RDS"))
}
#excess_birth_year_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_mun_df",".RDS"))
#excess_birth_year_adj2_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_adj2_mun_df",".RDS"))

#excess by citizenship region
excess_birth_ctzreg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_ctzreg_df",".RDS"))
excess_birth_year_ctzreg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_ctzreg_df",".RDS"))
#excess_birth_ctzreg_ctn_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_ctzreg_ctn_df",".RDS"))

#population (women 15-50) by country
pop_detctz_df = readRDS(paste0(code_root_path,"savepoint/","pop_detctz_country_df.RDS"))

#decile analysis
excess_by_ntiles_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_ntiles_df",".RDS"))
excess_by_ntiles_swiss_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","swiss","_","seedid",seed_id,"_","excess_birth_ntiles_df",".RDS"))
excess_by_ntiles_nonswiss_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","non-swiss","_","seedid",seed_id,"_","excess_birth_ntiles_df",".RDS"))
excess_by_ntiles_first_df = readRDS(paste0(code_root_path,"results/",save.date2,"_",mod_name,"_","first","_","seedid",seed_id,"_","excess_birth_ntiles_df",".RDS"))
excess_by_ntiles_second_df = readRDS(paste0(code_root_path,"results/",save.date2,"_",mod_name,"_","second","_","seedid",seed_id,"_","excess_birth_ntiles_df",".RDS"))
```

# Figures

::: {#fig-myfig}

```{r,  fig.height=6,fig.width=10}
p1 = excess_birth_year_month_df %>% 
  ggplot(aes(x=date))+
  geom_ribbon(aes(ymin=n_exp_lwb, ymax=n_exp_upb),alpha=0.2)+
  geom_line(aes(y=n_exp_mean)) +
  geom_point(aes(y=n_birth), alpha=0.6,col="darkred")+
  scale_x_date(name="", expand = c(0.01, 0.01),limits=dmy(c("01-01-2000","01-01-2025")))+
  scale_y_continuous(name="Monthly births")

#fertility
p2 = birth_prob_by_age_df %>% 
    filter(year %in% c(2000,2010,2024)) %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
    geom_line(aes(col=factor(year)))+
    scale_x_continuous(name="Maternal age")+
    scale_y_continuous(name="Probability of having a child\n(by month)",labels = scales::percent)+
    scale_color_discrete(name="Calendar year")+
    scale_fill_discrete(name="Calendar year")


#calculate peak age
df <- birth_prob_by_age_df %>% filter(year==2000) %>%  arrange(mother_age)
idx <- which.max(df$est)
y1 <- df$est[idx - 1]
y2 <- df$est[idx]
y3 <- df$est[idx + 1]
x0 <- df$mother_age[idx]
delta <- (y1 - y3) / (2 * (y1 - 2 * y2 + y3))
peak_age <- x0 + delta

p3 = gp_rel_df %>% 
  dplyr::mutate(est = est + peak_age,
                lwb = lwb + peak_age,
                upb = upb + peak_age) %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="black",alpha=0.2)+
  geom_line(col="black")+
  scale_y_continuous(name="Age of fertility peak")+
  scale_x_continuous(name="Calendar year")

cowplot::plot_grid(p1,
                   cowplot::plot_grid(p2,p3,rel_widths = c(1.4,1),ncol=2,labels = c("B.","C.")),
                   rel_heights = c(1,1),labels = c("A.",""),
                   nrow = 2)
```

**Model fit**. A. Monthly number of births, as observed (red points) and expected (black lines corresponding to the mean posterior and gray area to the 95% credibility interval (95% CrI)), from 2000 to 2024. B. Fertility curve. Probability of having a child (by month) according to maternal age for different calendar years (2000, 2010 and 2024). C. Age of fertility peak according to calendar year.
:::


::: {#fig-myfig}

```{r,  fig.height=8,fig.width=12}
#national, year
plot_df = excess_birth_year_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))

p1 = plot_df %>% 
  filter(var %in% c("Expected birth")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean)) +
  geom_point(aes(y=n_birth),col="darkred" ) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Annual birth")

p2 = plot_df %>% 
  filter(var %in% c("Relative excess")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean))+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)

#national, year, month
plot_df = excess_birth_year_month_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
p3 = plot_df %>% 
  filter(year>=2020, var=="Relative excess") %>% 
    ggplot(aes(x = date)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean))+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_date(name="")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)


#-------------------------------------------------------------------------------
rel_pop_detctz_df = pop_detctz_df %>% 
  left_join(ctz_map,by="ctz_name") %>% 
  group_by(ctz_region,year) %>% 
  dplyr::summarise(n=sum(n),.groups="drop_last") %>% 
  dplyr::summarise(n=mean(n),.groups="drop") %>% 
  dplyr::mutate(p=n/sum(n))
region_levels <- excess_birth_year_ctzreg_df %>%
  left_join(rel_pop_detctz_df, by="ctz_region") %>%
  distinct(ctz_region, p) %>%
  arrange(-p) %>%
  mutate(ctz_region2 = paste0(ctz_region, " (", round(p*100, 1), "%)")) %>%
  pull(ctz_region2)

n=15
cols <- scales::hue_pal()(n)
eu_colors <- cols[c(1,3,5)] #scales::seq_gradient_pal("darkblue", "#ABD9E9")(seq(0,1,length.out=3))
rest_colors <- cols[c(7,8,9,10,14,15)]# scales::seq_gradient_pal("#D7191C", "#FDAE61")(seq(0,1,length.out=6))

world_reg_name = c("Switzerland", "Western Europe", "Eastern Europe",
                   "South America", "North America", "Central America",
                   "Africa", "Asia", "Oceania")

rel_pop_detctz_df <- rel_pop_detctz_df %>%
  mutate(ctz_region2 = paste0(ctz_region, " (", round(p*100, 1), "%)"),
         ctz_region2 = factor(ctz_region2, levels = region_levels)) %>%
  #sort so that colors are in right orders
   mutate(ctz_region = factor(ctz_region, levels = world_reg_name),
          ctz_region3 = gsub(" ", "\n", ctz_region),
          ctz_region4 = gsub(" ", "\n", ctz_region2)) %>% 
  arrange(ctz_region)
rel_pop_detctz_df$fill_col = c(eu_colors,rest_colors)

p4 = excess_birth_ctzreg_df %>%
  left_join(rel_pop_detctz_df,by="ctz_region") %>% 
  ggplot(aes(x=ctz_region4,y=rel_exc_mean,ymin=rel_exc_lwb,ymax=rel_exc_upb,col=ctz_region4))+
  geom_hline(yintercept = 0, lty=2)+
  geom_pointrange()+
  scale_x_discrete(name="",
                   limits = rel_pop_detctz_df$ctz_region4)+
  scale_y_continuous(name = "Relative excess birth",breaks=c(-0.2,0,0.2,0.4,0.6,0.8,1),labels = scales::percent)+
  scale_color_manual( name = "Region",
                        values = rel_pop_detctz_df$fill_col,
                        breaks = rel_pop_detctz_df$ctz_region4) +
  theme(legend.position = "none")
p5 = excess_birth_year_ctzreg_df %>%
  left_join(rel_pop_detctz_df, by="ctz_region") %>%
  mutate(region_group = if_else(ctz_region %in% c("Switzerland", "Western Europe", "Eastern Europe"),
                                "Europe", "Other")) %>% 
  filter(p>0.01) %>% 
  ggplot(aes(x=year, y=rel_exc_mean, ymin=rel_exc_lwb, ymax=rel_exc_upb)) +
  geom_ribbon(aes(fill=ctz_region), alpha=0.1) +
  geom_line(aes(col=ctz_region)) +
  facet_wrap(region_group~.,nrow=2,scales="free_y") +
  geom_hline(yintercept = 0, lty=2) +
   scale_y_continuous(name = "Relative excess birth",breaks=c(-0.2,0,0.2,0.4,0.6,0.8,1),labels = scales::percent)+
  scale_x_continuous(name="",breaks=c(2011,2015,2020,2024))+
  scale_color_manual( name = "",
                        values = rel_pop_detctz_df$fill_col,
                        breaks = rel_pop_detctz_df$ctz_region)+
  scale_fill_manual( name = "",
                        values = rel_pop_detctz_df$fill_col,
                        breaks = rel_pop_detctz_df$ctz_region)+
  theme(legend.position = "bottom",
        legend.box.margin = margin(t = -18))


#plot
cowplot::plot_grid(cowplot::plot_grid(p1,
                                      p2,
                                      p3,
                                      nrow=3,labels=c("A.","B.","C.")),
                   cowplot::plot_grid(p4,
                                      p5+theme(legend.position = "none"),
                                      get_legend2(p5),
                                      nrow=3,labels=c("D.","E.",""),rel_heights = c(1,1.8,0.2)),
                   ncol=2,rel_widths = c(1.2,1))
```

**Excess birth**. A. Annual number of births, as observed (red points) and expected (black lines corresponding to the mean posterior and gray area to the 95% credibility interval (95% CrI)), from 2000 to 2024. B. Relative excess birth by year from 2000 to 2024. C. Relative excess birth by month between 2020 and 2024. D. Relative excess birth by citizenship aggregated by region of the world in 2011-2024. E. Relative excess birth by citizenship over years.
:::

::: {#fig-myfig}

```{r,  fig.height=8,fig.width=10}
excess_birth_adj2_mun_df %>% 
  #left_join(new_mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  left_join(new_mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>% 
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean), color = NA) +  # communes sans bordure
  geom_sf(data = regions_sf %>% mutate(dist_id = as.numeric(dist_id)), 
          fill = NA, color = "black", size = 0.3) +  # contours districts
  geom_sf(data = lake_sf, fill = "lightblue", color = NA, alpha = 0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",
    mid = "lightyellow",
    high = "green",
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1),
    #limits=c(-0.4,0.4)
  ) +
  theme(legend.position = "bottom",
        legend.direction = "horizontal")
```

**Excess birth by municipality**. Relative excess birth by municipality between 2017 and 2024.
:::



::: {#fig-myfig}

```{r,  fig.height=6,fig.width=10}
#plot
p1 = excess_by_ntiles_df %>% 
  left_join(data.frame(explanatory_var  = c("pop_dens_building_ntile", "rel_empl_mean_ntile", "sep_ntile","votation_paternity_leave_ntile"),
                       explanatory_var2 =  c("Population density", "Childcare service density","Socio-economic position (SEP)","Share of votes in favour of paternity leave extension")),by="explanatory_var") %>% 
   filter(!is.na(explanatory_var2)) %>% 
  dplyr::mutate(explanatory_var2 = factor(explanatory_var2, levels = c("Population density", "Childcare service density","Socio-economic position (SEP)","Share of votes in favour of paternity leave extension"))) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(alpha = 0.2) +
  geom_line() +
  geom_point() +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_x_continuous(name="Population decile",breaks=c(1:10))+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var2~.,ncol=2)
p1
```

**Relative excess birth stratified by variable deciles**
:::

::: {#fig-myfig}

```{r,  fig.height=8,fig.width=10}

subgroup_col = c("black", "darkred", "darkorange", "darkgreen","darkblue")
subgroup_names = c("all", "swiss","non-swiss","first","second")
subgroup_names2 = c("All", "Swiss","Non-Swiss","1st births","2nd+ births")


#birth prob
birth_prob_by_age_subgroup_df = rbind(birth_prob_by_age_df %>% filter(year==2000) %>% dplyr::mutate(subgroup="all"),
                                   birth_prob_by_age_swiss_df %>% filter(year==2000) %>% dplyr::mutate(subgroup="swiss"),
                                   birth_prob_by_age_nonswiss_df %>% filter(year==2000) %>% dplyr::mutate(subgroup="non-swiss"))
birth_prob_by_age_subgroup_df = birth_prob_by_age_subgroup_df %>% 
  left_join(birth_prob_by_age_subgroup_df %>% 
              filter(subgroup=="all") %>% 
              dplyr::select(mother_age,est_all = est),by="mother_age")

#excess
excess_birth_year_subgroup_df = rbind(excess_birth_year_df %>% dplyr::mutate(subgroup="all"),
                                      excess_birth_year_swiss_df  %>% dplyr::mutate(subgroup="swiss"),
                                      excess_birth_year_nonswiss_df  %>% dplyr::mutate(subgroup="non-swiss"),
                                      excess_birth_year_first_df  %>% dplyr::mutate(subgroup="first"),
                                      excess_birth_year_second_df  %>% dplyr::mutate(subgroup="second"))
excess_birth_year_subgroup_df = excess_birth_year_subgroup_df %>% 
  left_join(excess_birth_year_subgroup_df %>% 
              filter(subgroup=="all") %>% 
              dplyr::select(year,rel_exc_mean_all = rel_exc_mean),by="year")


#excess
excess_by_ntiles_subgroup_df = rbind(excess_by_ntiles_df %>% dplyr::mutate(subgroup="all"),
                                      excess_by_ntiles_swiss_df  %>% dplyr::mutate(subgroup="swiss"),
                                      excess_by_ntiles_nonswiss_df  %>% dplyr::mutate(subgroup="non-swiss"),
                                      excess_by_ntiles_first_df %>% filter(childless==TRUE)  %>% dplyr::mutate(subgroup="first"),
                                      excess_by_ntiles_second_df  %>% filter(childless==TRUE)  %>% dplyr::mutate(subgroup="second"))
excess_by_ntiles_subgroup_df = excess_by_ntiles_subgroup_df %>% 
  left_join(excess_by_ntiles_subgroup_df %>% 
              filter(subgroup=="all") %>% 
              dplyr::select(ntile,explanatory_var,rel_exc_mean_all = rel_exc_mean),by=c("ntile","explanatory_var"))

#plot
p1 = birth_prob_by_age_subgroup_df %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(subgroup)),alpha=0.2)+
    geom_line(aes(col=factor(subgroup)))+
    scale_x_continuous(name="Maternal age")+
    scale_y_continuous(name="Probability of having a child\n(by month)", labels = scales::percent)+
    scale_color_manual(name = "",
                       values = subgroup_col,
                       breaks =subgroup_names)+
    scale_fill_manual(name = "",
                      values = subgroup_col,
                       breaks =subgroup_names)+
  theme(legend.position = "none")

p2 = excess_birth_year_subgroup_df %>% 
  filter(subgroup %in% c("swiss","non-swiss")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = rel_exc_lwb ,ymax=rel_exc_upb ,fill=subgroup),alpha = 0.2) +
  geom_line(aes(y = rel_exc_mean , col=subgroup))+
  geom_line(aes(y = rel_exc_mean_all),linetype=3, linewidth = 2)+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)+
  scale_color_manual(name = "",
                       values = subgroup_col,
                       breaks =subgroup_names)+
    scale_fill_manual(name = "",
                      values = subgroup_col,
                       breaks =subgroup_names)+
   theme(legend.position = "none")

p3 = excess_by_ntiles_subgroup_df %>% 
  filter(subgroup %in% c("swiss","non-swiss")) %>% 
  left_join(data.frame(explanatory_var  = c("pop_dens_building_ntile"),
                       explanatory_var2 =  c("Population density")),by="explanatory_var") %>% 
   filter(!is.na(explanatory_var2)) %>% 
  dplyr::mutate(explanatory_var2 = factor(explanatory_var2, levels = c("Population density"))) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(aes(fill=subgroup),alpha = 0.2) +
  geom_line(aes(col=subgroup)) +
  geom_point(aes(col=subgroup)) +
  geom_line(aes(y = rel_exc_mean_all),linetype=3, linewidth = 2)+
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_x_continuous(name="Population decile",breaks=c(1:10))+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var2~.,ncol=2)+
    scale_color_manual(name = "",
                       values = subgroup_col,
                       breaks =subgroup_names)+
    scale_fill_manual(name = "",
                      values = subgroup_col,
                       breaks =subgroup_names)+
   theme(legend.position = "none")

p4 = excess_birth_year_subgroup_df %>% 
  filter(subgroup %in% c("first","second")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = rel_exc_lwb ,ymax=rel_exc_upb ,fill=subgroup),alpha = 0.2) +
  geom_line(aes(y = rel_exc_mean , col=subgroup))+
  geom_line(aes(y = rel_exc_mean_all),linetype=3, linewidth = 2)+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)+
    scale_color_manual(name = "",
                       values = subgroup_col,
                       breaks =subgroup_names)+
    scale_fill_manual(name = "",
                      values = subgroup_col,
                       breaks =subgroup_names)+
   theme(legend.position = "none")
  

p5 = excess_by_ntiles_subgroup_df %>% 
   filter(subgroup %in% c("first","second")) %>% 
  left_join(data.frame(explanatory_var  = c("pop_dens_building_ntile"),
                       explanatory_var2 =  c("Population density")),by="explanatory_var") %>% 
   filter(!is.na(explanatory_var2)) %>% 
  dplyr::mutate(explanatory_var2 = factor(explanatory_var2, levels = c("Population density"))) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(aes(fill=subgroup),alpha = 0.2) +
  geom_line(aes(col=subgroup)) +
  geom_point(aes(col=subgroup)) +
  geom_line(aes(y = rel_exc_mean_all),linetype=3, linewidth = 2)+
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_x_continuous(name="Population decile",breaks=c(1:10))+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var2~.,ncol=2)+
    scale_color_manual(name = "",
                       values = subgroup_col,
                       breaks =subgroup_names)+
    scale_fill_manual(name = "",
                      values = subgroup_col,
                       breaks =subgroup_names)+
   theme(legend.position = "none")

legend_plot = get_legend2(data.frame(year=2000:2001,
           mean = 1:2,
           lwb = 1:2,
           upb = 1:2) %>% 
   crossing(subgroup = subgroup_names) %>% 
  ggplot(aes(x = year, y = mean, ymin=lwb, ymax = upb))+
  geom_ribbon(aes(fill=subgroup),alpha = 0.2) +
  geom_line(aes(col=subgroup)) +
  geom_point(aes(col=subgroup)) +
  scale_color_manual(name = "",
                     labels = subgroup_names2,
                       values = subgroup_col,
                       breaks =subgroup_names,
                     guide = guide_legend(nrow = 3))+
    scale_fill_manual(name = "",
                      labels = subgroup_names2,
                      values = subgroup_col,
                       breaks =subgroup_names,guide = guide_legend(nrow = 3))+
    theme(legend.position = "bottom"))


cowplot::plot_grid(cowplot::plot_grid(p1,p2,p4,
                                      labels = c("A.","B.","D."),nrow=3,rel_heights = c(1,1,1)),
                   cowplot::plot_grid(legend_plot,p3,p5,
                                      labels = c("","C.","E."),nrow=3,rel_heights = c(1,1,1)),
                   ncol=2,rel_widths = c(1,1))

```

**Sub-group excess births**: 
:::




```{r, include=FALSE}

#graph to compare fertility curves including any child or first child only
if(FALSE){
  supp_p1 = gamma_month_df %>% 
  ggplot(aes(x=month_id,y=est,ymin=lwb,ymax=upb)) +
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")
  
  birth_prob_by_age_df1=  readRDS(paste0(code_root_path,"results/","20260309","_","mod8","_seedid",1,"_birthprob.RDS"))
  birth_prob_by_age_df2=  readRDS(paste0(code_root_path,"results/","20260320","_","mod8_first","_seedid",1,"_birthprob.RDS"))
  rbind( birth_prob_by_age_df1 %>% dplyr::mutate(data="all"),
         birth_prob_by_age_df2 %>% dplyr::mutate(data="first")) %>% 
    filter(year %in% c(2024)) %>% 
    #group_by(data) %>% dplyr::mutate(lwb=lwb/max(est),upb=upb/max(est),est=est/max(est)) %>% ungroup() %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(data)),alpha=0.2)+
    geom_line(aes(col=factor(data)))+
    scale_y_continuous(labels = scales::percent)
}
```