---
title: "Birth decline in Switzerland"
format:
  html:
    theme: cosmo
    code-fold: true        # adds "Show/Hide Code" buttons
    code-summary: "Code"   # customize button text
    toc: true
    toc-location: left  # place TOC on the left
    number-sections: true  # number sections
    toc-depth: 3      # show subsections up to 3 levels
    toc-fold: true     # allow collapsing/expanding sections
execute:
  fig-width: 7
  fig-height: 5
  warning: false
  message: true
  comment: ""
editor: visual
---

```{r,include=FALSE}
source("../R/000_setup.R")

#load individual birth data, 1987_2024
#birth_df = load_birth_data(mun_df = mun_df, rerun=FALSE)
#load cleaned, aggregated data
load(paste0(code_root_path,"savepoint/cleaned_df.RData"))
```

# Objectives

- Quantify Switzerland’s birth deficit over 2000–2024, adjusting for demographic changes and delayed childbearing
- Examine temporal and geographical heterogeneity in birth deficits
- Explore socio-economic and policy factors associated with variations in excess/deficit births

# Data

## Types of data

- Birth data:
  - Individual-level: year, month, municipality, maternal age  
  - Period: 1987–2024 (analysis uses 2000–2024)  

- Population data:
  - Aggregated: year × canton, nationality (Swiss vs non-Swiss), 1987–2024  
  - Optionally: municipality-level counts 2011–2024  

- Socio-economic data (Explanatory variables)
  - Votation data (municipality-level):
    - family/education support
    - paternity leave
    - childcare tax reductions
    - women’s suffrage  
  - Swiss Household data (canton-level)  
  - Socio-economic position (municipality-level)  
  
## Birth data

```{r,fig.width=6,fig.height=8}
birth_df1 = birth_agg_df %>% 
  filter(mother_age %in% 15:50,year>=2000) %>% 
  dplyr::mutate(age_group = cut(mother_age,c(15,25,35,Inf),  right = FALSE, labels = c("15-24", "25-34", "35-50"))) %>% 
  group_by(year,age_group) %>% 
  dplyr::summarise(n=sum(n), .groups="drop")

birth_df2 = rbind(birth_df1 %>% dplyr::mutate(data="by age"),
                  birth_df1 %>% 
                    group_by(year)  %>% 
                    dplyr::summarise(n=sum(n),.groups="drop") %>% 
                    dplyr::mutate(age_group="All ages",data="All ages"))


pop_df1 = pop_df %>% 
  filter(age %in% 15:50,year>=2000) %>% 
  dplyr::mutate(age_group = cut(age,c(15,25,35,Inf),  right = FALSE, labels = c("15-24", "25-34", "35-50"))) %>% 
  group_by(year,age_group) %>% 
  dplyr::summarise(n=sum(n), .groups="drop")

pop_df2 = rbind(pop_df1 %>% dplyr::mutate(data="by age"),
                pop_df1 %>% 
                    group_by(year)  %>% 
                    dplyr::summarise(n=sum(n),.groups="drop") %>% 
                    dplyr::mutate(age_group="All ages",data="All ages"))


scientific_10 <- function(x) {   ifelse(x==0,0,parse(text=gsub("e\\+*", " %*% 10^", scales::scientific_format()(x)))) }
p1 = birth_df1 %>% 
  dplyr::mutate(year = dmy(paste0("01-01-", year))) %>% 
  ggplot(aes(x=year, y = n))+
  geom_bar(stat="identity",alpha=0.5)+
  scale_x_date(name="")+
  scale_y_continuous(name="Number of births")
p2 = pop_df1 %>% 
  dplyr::mutate(year = dmy(paste0("01-01-", year))) %>% 
  ggplot(aes(x=year, y = n))+
  geom_bar(stat="identity",alpha=0.5)+
  scale_x_date(name="")+
  scale_y_continuous(name="Population size", label=scientific_10)

p3 = birth_df2 %>% 
  rename(n_birth = n) %>% 
  left_join(pop_df2 %>% rename(n_pop = n),
            by = c("year","age_group")) %>% 
  dplyr::mutate(p_birth = n_birth / n_pop,
                year = dmy(paste0("01-01-", year))) %>% 
  ggplot(aes(x = year, y = p_birth,
             colour = age_group,
             linetype = age_group)) +
  geom_line(,linewidth = 1) +
  expand_limits(y = 0)+
  scale_x_date(name="")+
  scale_y_continuous(name="Birth rate",labels = scales::percent)+
  scale_color_manual(name="Age groups",values=c("darkred","orange","darkblue","black"))+
  scale_linetype_manual(name="Age groups",values = c("dashed","dashed","dashed","solid"))+
  theme(legend.position = "bottom")

cowplot::plot_grid(p1,
                   p2,
                   p3, ncol=1,rel_heights = c(1,1,1.2),labels = c("A.","B.","C."))

```

**Figure 1**: A. Number of births by year. B. Female population aged 15-50 by year. C. Birth rates by year and age group.

# Methods

## Birth prediction model

We model age-specific births using a negative binomial regression with an overdispersion parameter. Let:

- $y_{a,t}$ = number of births at maternal age $a$ in year $t$  
- $N_{a,t}$ = population of women at age $a$ in year $t$  

Then:

$$
y_{a,t} \sim \text{NegBinom}(\mu_{a,t}, \phi), \quad 
\log\mu_{a,t} = f(a + \delta_t) + \gamma_m + \log N_{a,t},
$$

where:

- $f(a)$ is the **age-specific fertility function** modeled with a Gaussian Process (GP),
- $\delta_t$ is the **calendar-year-specific age shift** capturing delayed childbearing (also modeled with a GP),
- $\gamma_m$ is a **fixed effect for calendar month** $m$,
- $\phi$ is the negative binomial **overdispersion parameter**.  

This formulation allows the fertility curve to shift over time while accounting for seasonality and overdispersion in birth counts.

## Excess births

### National-level

Let $\hat{\mu}_{a,t}$ = posterior predicted births from the model, the excess is:

$$
\text{Excess}_{a,t} = y_{a,t} - \hat{\mu}_{a,t}, \quad
\text{Relative Excess}_{a,t} = \frac{y_{a,t} - \hat{\mu}_{a,t}}{\hat{\mu}_{a,t}}.
$$

This gives the absolute and relative deviations between observed and expected births.

### District-level

Let $d$ denote a district and $y_{d,a,t}$ the observed births in district $d$ (unknown at prediction stage).
We distribute the national posterior predicted births $\hat{\mu}_{a,t}$ across districts using a **multinomial distribution**:

$$
\hat{\mu}_{1\cdots D,a,t} \sim \text{Multinomial}\Big(\hat{\mu}_{a,t},p_{1\cdots D,a,t} \Big).
$$

Where $p_{d,a,t}$ is the proportion of women of age $a$ in district $d$ relative to the total national population of age $a$:

$$
p_{d,a,t} = \frac{N_{d,a,t}}{\sum_{d'} N_{d',a,t}}.
$$

The district-level excess is then:

$$
\text{Excess}_{d,a,t} = y_{d,a,t} - \hat{\mu}_{d,a,t}, \quad
\text{Relative Excess}_{d,a,t} = \frac{y_{d,a,t} - \hat{\mu}_{d,a,t}}{\hat{\mu}_{d,a,t}}.
$$

This allows us to quantify deviations at a finer spatial scale while respecting the national posterior predictions.


## Explaining heterogeneity

- Temporal and geographical differences analyzed with covariates  
- Focus on structural factors beyond age composition  

# Results: excess

## Model parameters

```{r}
#load data
save.date="20260214"
mod_name = "mod5"
seed_id=8

#load results
par_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_par.RDS"))
birth_prob_by_age_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_birthprob.RDS"))
gp_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gp.RDS"))
gp_rel_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gp_rel.RDS"))
age_bias_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_agebias.RDS"))
pred_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_pred.RDS"))
par_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_par.RDS"))

#load excess
excess_birth_year_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_age_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_df",".RDS"))
excess_birth_year_month_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_df",".RDS"))
excess_birth_year_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_reg_df",".RDS"))

excess_birth_year_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_ctz_df",".RDS"))
excess_birth_year_age_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_ctz_df",".RDS"))
excess_birth_year_month_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_ctz_df",".RDS"))
```

### Month effect

```{r}
par_df %>%
  filter(str_detect(variable, "gamma_month")) %>% 
  tidyr::extract(variable,into=c("variable","month_id"),
                 regex =paste0('(\\w.*)\\[',paste(rep("(.*)",1),collapse='\\,'),'\\]'), remove = T) %>%
  as_tibble() %>%
  dplyr::select(month_id,est=mean,lwb=`2.5%`,upb=`97.5%`) %>%
  dplyr::mutate(month_id=as.numeric(month_id)) %>% 
  ggplot(aes(x=month_id,y=est,ymin=lwb,ymax=upb)) +
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")
```

### Fertility curve

```{r}
birth_prob_by_age_df %>% 
    filter(year %in% c(2000,2010,2024)) %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
    geom_line(aes(col=factor(year)))
```

### Age shift

```{r}
gp_rel_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")+
  scale_y_continuous(name="Change in age of fertility peak (relative to 2000)")
```

## Excess mortality

### National level

```{r}
#national, year
plot_df = excess_birth_year_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~.,scale="free_y")
```

### Monthly level (2020-24)

```{r}
#national, year, month
plot_df = excess_birth_year_month_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year>=2020) %>% 
  ggplot(aes(x = date, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>%  filter(year>=2020) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>%  filter(year>=2020), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~.,scale="free_y")

```

### District-level

```{r, fig.height=16,fig.width=12}
#load region location
regions_sf <- st_read(paste0(code_root_path,"data/boundary_data/Boundaries_K4_District_20260101.shp")) #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
regions_sf = regions_sf[,c(1,3,7)]
names(regions_sf)[c(1,2)] <- c("dist_name2", "dist_id")

excess_birth_year_reg_df %>% 
  filter(year %in% 2017:2024) %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = rel_exc_mean)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(year~.,ncol=2)

```

### Municipality-level

#### By municipality, by year

```{r, fig.height=16,fig.width=12}
#load region location
excess_birth_year_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_mun_df",".RDS"))
mun_sf <- st_read(paste0(code_root_path,"data/boundary_data/Boundaries_K4_Commune_20260101.shp")) #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
mun_sf = mun_sf[,c(1,3,7)]
names(mun_sf)[c(1,2)] <- c("mun_name2", "mun_id")

lake_sf <- st_read(paste0(code_root_path,"data/boundary_data/swissTLMRegio_Product_LV95/Hydrography/swissTLMRegio_Lake.shp"))
mun_buffer <- st_buffer(mun_sf, dist = 1000)  # distance in CRS units (LV95 ~ meters)
# keep lakes that intersect any municipality buffer
lake_sf <- lake_sf %>% 
  st_filter(mun_buffer, .predicate = st_intersects)

excess_birth_year_mun_df %>% 
  filter(year %in% 2017:2024) %>% 
  dplyr::mutate(rel_exc_mean = case_when(rel_exc_mean>1 ~1,
                                         TRUE ~ rel_exc_mean)) %>% 
  left_join(mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean),color = "black") +
  geom_sf(data=lake_sf,fill = "lightblue",color=NA,alpha=0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(year~.,ncol=2)

```

#### Summed over adjacent municipality, by year

```{r, fig.height=16,fig.width=12}
#load region location
excess_birth_year_adj_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_adj_mun_df",".RDS"))

excess_birth_year_adj_mun_df %>% 
  filter(year %in% 2017:2024) %>% 
  dplyr::mutate(rel_exc_mean = case_when(rel_exc_lwb >0 | rel_exc_upb<0 ~ rel_exc_mean ,
                                         TRUE ~ 0)) %>% 
  left_join(mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  #left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  left_join(mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean),color = "black",size = 0.1) +
  geom_sf(data=lake_sf,fill = "lightblue",color=NA,alpha=0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(year~.,ncol=2)

```


#### Summed over adjacent municipalities, 2017-2024

```{r, fig.height=6,fig.width=8}
#load region location
excess_birth_adj_mun_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_adj_mun_df",".RDS"))

excess_birth_adj_mun_df %>% 
  dplyr::mutate(rel_exc_mean = case_when(rel_exc_lwb >0 | rel_exc_upb<0 ~ rel_exc_mean ,
                                         TRUE ~ 0)) %>% 
  left_join(mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  #left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  left_join(mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean),color = "black",size = 0.1) +
  geom_sf(data=lake_sf,fill = "lightblue",color=NA,alpha=0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")
```


# Data: explanatory variables

## Votation data

```{r,fig.width=8,fig.height=8}
#load region location
regions_sf <- st_read(paste0(code_root_path,"data/boundary_data/Boundaries_K4_District_20260101.shp")) #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
regions_sf = regions_sf[,c(1,3,7)]
names(regions_sf)[c(1,2)] <- c("dist_name2", "dist_id")

vote_dist_df %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = yes_prop/100)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0.5,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(vote_object~.,ncol=2)
```

**Figure 2**: Votation results by municipality for the four objects.

## Rural/Urban data

```{r,fig.width=6,fig.height=6}
rural_urban_dist_df %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = urban_rural_type2)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Rural level",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 20
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")
```

```{r,fig.width=6,fig.height=6}
rural_urban_df %>% 
   left_join(mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>%
  dplyr::mutate(urban_rural_type2 =  dense_rank(urban_rural_type2)) %>% 
  st_as_sf() %>%
  ggplot(aes(fill = urban_rural_type2)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Rural level",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 5
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")
```

**Figure XX**: Level of rurality.

## Swiss Household Panel (SHP) data

```{r,fig.width=10,fig.height=10}
p_df = readRDS(paste0(code_root_path,"savepoint/cleaned_shp_p_df.RDS"))

#variables
  #personal data
  variables_p_df = data.frame(var= c("idhous","year", "idpers", "filter","pdate", "status","plingu",#"idint",
                                     "age","sex",
                                     "pd80", "pd81", "pd82",
                                     "pd90", "pd91", "pd92", "pd93", "pd94", "pd95",
                                     "pp20", "pp21", "pp22", "pp23"),
                              def =  c("Identification number of household",
                                       "Year",
                                       "identification number of the person",
                                       "Identification of the survey",
                                       "Date of personal interview",
                                       "Type of interviews completed: grid, proxy, personal",
                                       "Interview language",# "Identification number of interviewer",

                                       "Age in year of interview",
                                       "Sex",

                                       "Total number children wanted",
                                       "Child wanted in the next 24 months",
                                       "Ideal family size",

                                       "Opinion on family: marriage is an outdated institution",
                                       "Opinion on family: job preserves independance",
                                       "Opinion on family: child suffers with working mother",
                                       "Opinion on family: child suffers with unmaried parents",
                                       "Opinion on family: happiness of a child living with a single parent",
                                       "Opinion on family: happiness of a child living with same-sex parents",

                                       "Gender: Women in general penalized",
                                       "Gender: Personnally penalized",
                                       "Gender: In favour of measures",
                                       "Gender: Personal action" ))
  #household data
  variables_h_df = data.frame(var= c("idhous","year", "refper", "filter", "hmode","hdate","stathh","sthhre",#"idint"
                                     "nbkid",
                                     "canton",
                                     "ctn_abbr",
                                     "hf52"),
                              def =  c("Identification number of household",
                                       "Year",
                                       "Identification of reference person",
                                       "Identification of the survey",
                                       "Mode of data collection",
                                       "Date of household interview with reference person",
                                       "Status household questionnaire",
                                       "Reason for not completing household questionnaire",#"Identification number of interviewer",
                                       "Number of children in household: 0 to 17 years",
                                       "Canton of residence",
                                       "Canton of residence (abbr)",
                                       "External help: Child care: Monthly expenses"))
  #all variables
  variables_df = rbind(variables_p_df,
                       variables_h_df %>% filter(!(var %in% variables_p_df$var))) %>%
    dplyr::mutate(group = dplyr::case_when(
                            var %in% c("idhous", "idpers", "refper") ~ "identifiers",
                            var %in% c("year", "pdate", "hdate") ~ "time",
                            var %in% c("status", "hmode", "filter", "stathh", "sthhre") ~ "survey_process",
                            var %in% c("age", "sex", "plingu", "canton") ~ "sociodemographics",
                            var %in% c("pd80","nbkid") ~ "number_of_children",
                            var %in% c("pd81") ~ "child_wanted",
                            grepl("^pd", var) ~ "family_opinions_preferences",
                            grepl("^pp", var) ~ "gender_attitudes",
                            var %in% c("hf52") ~ "household_children_economics",
                            TRUE ~ NA )) %>%
    dplyr::mutate(is_reg_var = ifelse(group %in% c("family_opinions_preferences","gender_attitudes","household_children_economics",
                                                   "number_of_children","child_wanted"), 1, 0),
                 type = case_when(var %in% c("pd81") ~ "factor_2lev_n_y",
                                   var %in% c("pd90", "pd91", "pd92", "pd93", "pp20", "pp21", "pp22", "pp23")  ~  "factor_11lev_0_10",
                                   var %in% c("pd80", "pd82","nbkid")  ~ "numeric_max10",
                                   var %in% c("pd94", "pd95")  ~ "factor_11lev_0_10",
                                   var %in% c("hf52") ~ "numeric",
                                   TRUE ~ NA))

  df = p_df %>%
    dplyr::mutate(dplyr::across(everything(), as.character)) %>%
    dplyr::mutate(row_id = dplyr::row_number()) %>%
    tidyr::pivot_longer(cols = -c(year,row_id), names_to = "var", values_to = "value") %>%
    dplyr::group_by(var,year) %>%
    dplyr::summarise(p_na = mean(is.na(value)),
                     n_tot = n(),
                     n_nna = sum(!is.na(value)), .groups = "drop") %>%
    left_join(variables_df,by="var") %>%
    filter(is_reg_var==1) %>%
    dplyr::mutate(year = ymd(paste0(year,"-01-01")))
  
  df %>%
    ggplot(aes(x=year,y=n_nna))+
    geom_line(aes(x=year,y=n_tot),col="black",linetype = 2)+
    geom_point(aes(col=var))+geom_line(aes(col=var))+
    scale_y_continuous(name="Number of non-missing")+
    scale_color_discrete(breaks=variables_df$var,labels=variables_df$def,
                         guide = guide_legend(ncol = 2)) +
    expand_limits(y=0)+
    facet_wrap(group~.)+
    theme(legend.position = "bottom")
  
```

**Figure XX**: Number of responses.

```{r,fig.width=10,fig.height=10}
  df %>%
    ggplot(aes(x=year,y=1-p_na,col=var))+
    geom_point()+geom_line()+
    scale_y_continuous(labels = scales::percent,name="Proportion of non-missing")+
    scale_color_discrete(breaks=variables_df$var,labels=variables_df$def,
                         guide = guide_legend(ncol = 2)) +
    expand_limits(y=0)+
    facet_wrap(group~.)+
    theme(legend.position = "bottom")
```

**Figure XX**: Proportion of non-missing responses.

```{r,fig.width=8,fig.height=4}
  #number of new participants per year
  p_df %>%
    dplyr::mutate(is_woman_15_50 = factor(as.numeric(sex=="woman" & age>=15 & age<=50))) %>% 
    group_by(year_start,is_woman_15_50) %>%
    dplyr::summarise(n_new = length(unique(idpers)),.groups="drop") %>%
    ggplot(aes(x=year_start,y=n_new,fill=is_woman_15_50))+
    geom_bar(stat="identity",position = "stack")+
    scale_y_continuous(name="Number of new participants")+
    scale_fill_manual(name="Participants",values=c("gray","blue"),breaks = c(0,1),labels=c("All","Woman 15-50"))

```

**Figure XX**: Number of new participants per year

```{r,fig.width=10,fig.height=16}
  #mean
  df = p_df %>%
    filter(year_start %in% c(2010:2024),sex=="woman",age_start %in% 15:50) %>%
    dplyr::mutate(dplyr::across(everything(), as.character)) %>%
    dplyr::mutate(row_id = dplyr::row_number()) %>%
    tidyr::pivot_longer(cols = -c(year,row_id), names_to = "var", values_to = "value") %>%
    left_join(variables_df,by="var") %>%
    filter(is_reg_var==1,!is.na(value)) %>%
    dplyr::mutate(value=as.numeric(value)) %>%
    group_by(year,var,def,group) %>%
    dplyr::summarise(mean = mean(value),
                     n_values = n(),.groups="drop") %>%
    dplyr::mutate(year = ymd(paste0(year,"-01-01"))) %>%
    group_by(year,group) %>%
    dplyr::mutate(n_values_min = min(n_values),
                  n_values_max = max(n_values),
                  ymax=max(mean)*1.1) %>% ungroup() %>%
    group_by(group) %>%
    dplyr::mutate(ymax=max(mean)*1.1) %>% ungroup()

  plots <- df %>%
    filter(n_values>=10) %>%
    dplyr::mutate(def = str_wrap(def, width = 20)) %>%
    split(.$group) %>%
    map(~ ggplot(.x, aes(x = year, y = mean)) +
          geom_point(aes(col = def)) + geom_line(aes(col = def)) +
          geom_text(aes(y = 0, label = n_values_min),
                    vjust = 0, show.legend = FALSE) +
          geom_text(aes(y = ymax, label = n_values_max),
                    vjust = 0, show.legend = FALSE) +
          expand_limits(y = 0) +
          expand_limits(y = max(.x$ymax*1.05, na.rm = TRUE))+
          ggtitle(unique(.x$group)) +
          theme(legend.position = "bottom"))
  cowplot::plot_grid(plotlist = plots, ncol = 1)
```

**Figure XX**: Responses from female participants aged 15-50.

```{r,fig.width=8,fig.height=6}
  p_df %>%
    filter(year_start %in% c(2010:2015),sex=="woman",age_start %in% 15:50) %>%
    dplyr::select(idhous,idpers,year,year_start,plingu,age,ctn_abbr,pp21) %>%
    left_join(canton_df %>% dplyr::select(ctn_abbr=ctn,NUTS2_name),by="ctn_abbr") %>%
    filter(!is.na(pp21),!is.na(ctn_abbr)) %>%
    group_by(year,NUTS2_name) %>%
    dplyr::summarise(mean = mean(pp21),
                     n_values = n()) %>% ungroup() %>%
    dplyr::mutate(year = ymd(paste0(year,"-01-01"))) %>%
    ggplot(aes(x = year, y = mean)) +
    geom_point(aes(col = NUTS2_name)) + geom_line(aes(col = NUTS2_name)) +
    geom_text(aes(y = mean, label = n_values),
              vjust = 0, show.legend = FALSE) +
    scale_y_continuous(name="Gender: Personnally penalized")+
    expand_limits(y = 0)
```

**Figure XX**: Variation over regions.







# To do

- SHP data: explore
- Municipality level (by year): 25 * 2154: cluster/hotspot
- urbanisation level

# Conclusion

- Persistent birth deficit in Switzerland after accounting for demographic shifts and delayed motherhood  
- Stronger deficit in urban areas and among Swiss residents  
- Monthly analysis suggests short-term pandemic effects  
- Next steps: link heterogeneity to socio-economic and policy factors