---
title: "Birth decline in Switzerland: "
format: html
editor: visual
execute:
  warning: false
  message: false
  echo: false
---

# Data

```{r}
#| include: false
source("../R/000_setup.R")

#load individual birth data, 1987_2024
birth_df = load_birth_data(mun_df = mun_df, rerun=FALSE)
#load cleaned, aggregated data
load(paste0(code_root_path,"savepoint/cleaned_df.RData"))

birth_df = birth_df %>% 
  dplyr::mutate(mother_birth_year = year - mother_age,)
```

## Population data

Stratification level: age, year (1987-2024), canton

-   Restrict to women between 15 and 50.
-   Linear interpolation within months
-   Finer level of stratification (e.g., municipality) for 2010-2024

### Age distribution by citizenship status

```{r}
d1 = pop_mun_df %>%
  filter(year %in% c(2011,2015,2020,2024)) %>% 
  group_by(year,citizenship,age) %>% 
  dplyr::summarise(n=sum(n),.groups = "drop_last") %>% 
  dplyr::mutate(p=n/sum(n)) %>% ungroup()
d2 = pop_mun_df %>%
  filter(year %in% c(2011,2015,2020,2024)) %>% 
  group_by(year,age) %>% 
  dplyr::summarise(n=sum(n),.groups = "drop_last") %>% 
  dplyr::mutate(p=n/sum(n)) %>% ungroup()
d1 %>% 
  ggplot(aes(x=age,y=p))+
  geom_bar(aes(fill=citizenship),stat = "identity",position = position_dodge(width=1))+
  geom_line(data=d2,col="blue")+
  scale_y_continuous(labels = scales::percent)+
  facet_grid(year~.)
```

## Birth data

Stratification level: individual-based

-   increase in high age and decrease in low age over time
-   probability of giving birth by age over time
-   municipality

### Number of births

#### Over time

```{r}
#Total number of births---------------------------------------------------------
#by year: decrease from 2016, exception in 2021
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n)) + geom_point()+geom_line()+
  expand_limits(y = 0)

```

#### By month

```{r}
birth_df %>% 
  filter(year %in% c(1990,2000,2010,2020,2024)) %>% 
  group_by(month,year) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=month,y=n,col=factor(year))) + geom_point()+geom_line()
```

#### By citizenship

```{r}
#number of births by mother citizenship
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship==8100)) %>%
  group_by(year,mother_citizenship) %>% dplyr::summarise(n=n()) %>% 
  ggplot(aes(x=year,y=n,col=factor(mother_citizenship)))+geom_point()+geom_line()+
  expand_limits(y = 0)
```

### Mother age

#### Mean mother age over time

```{r}
birth_df %>% 
  group_by(year) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```

#### Mother age distribution by calendar year

```{r}
birth_df |>
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Year")
```

#### Mean mother age by citizenship

```{r}
birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  group_by(year, mother_citizenship) %>% 
  dplyr::summarise(mean = mean(mother_age)) %>% 
  ggplot(aes(x = year, y = mean, col = factor(mother_citizenship))) +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
```

#### Mother age distribution by calendar year and citizenship

```{r}
#| fig-width: 8
#| fig-height: 10
birth_df |>
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(year %in% c(1990, 2000, 2010, 2020, 2024)) |>
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_citizenship))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Proportion", color = "Swiss") +
  facet_grid(year ~ .)+
  theme(legend.position = "bottom")
```

#### Mother age by mother birth year (right-censoring adjusted)

```{r}
years <- c(1970, 1975, 1980, 1985, 1990, 1995)

birth_df %>% 
  filter(mother_birth_year %in% years,
         mother_age > (1987 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year")
```

#### Mother age (first child only, density)

```{r}
years <- c(1982, 1983, 1988, 1990)

birth_df %>% 
  dplyr::mutate(mother_citizenship = as.numeric(mother_citizenship == 8100)) %>%
  filter(parity == 1,
         mother_birth_year %in% years,
         mother_age > (2004 - min(years)),
         mother_age < (2024 - max(years))) %>% 
  ggplot(aes(x = mother_age,
             y = after_stat(density),
             color = factor(mother_birth_year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Mother age", y = "Dist", color = "Birth year") +
  facet_grid(mother_citizenship ~ .)
```

### Parity

#### Parity distribution over selected years

```{r}
birth_df %>% 
  filter(year %in% c(2005, 2010, 2020, 2024),
         parity <= 5) %>% 
  ggplot(aes(x = parity,
             y = after_stat(density),
             color = factor(year))) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  labs(x = "Parity", y = "Count", color = "Year")
```

#### Proportion of parity 2–4 over time

```{r}
birth_df %>% 
  filter(year >= 2005) %>% 
  group_by(year, parity) %>% 
  dplyr::summarise(n = n(), .groups = "drop_last") %>% 
  dplyr::mutate(p = n / sum(n)) %>% 
  ungroup() %>% 
  filter(parity >= 2, parity <= 4) %>% 
  ggplot(aes(x = year,
             y = p,
             col = factor(parity))) +
  geom_line() +
  geom_point() +
  expand_limits(y = 0)
```

## Votation data

```{r,fig.height=10,fig.width=10}
#load region location
regions_sf <- st_read(paste0(code_root_path,"data/boundary_data/Boundaries_K4_District_20260101.shp")) #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
regions_sf = regions_sf[,c(1,3,7)]
names(regions_sf)[c(1,2)] <- c("dist_name2", "dist_id")

vote_dist_df %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = yes_prop/100)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0.5,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(vote_object~.,ncol=2)
```

## Rural/Urban data

```{r,fig.height=6,fig.width=10}
rural_urban_dist_df %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = urban_rural_type2)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Rural level",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 20
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")
```

# Models

## Model 5

By age, year and months: $12\times 25 \times 36  = 10,800$ rows. Seasonal variation captured by a fixed month effect.

### Results: basic

```{r}
#load data
save.date="20260214"
mod_name = "mod5"
seed_id=8

excess_birth_year_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_df",".RDS"))
excess_birth_year_age_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_df",".RDS"))
excess_birth_year_month_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_df",".RDS"))
#excess_birth_year_month_age_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_age_df",".RDS"))
excess_birth_year_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_reg_df",".RDS"))
#excess_birth_year_age_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_reg_df",".RDS"))
#excess_birth_year_month_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_reg_df",".RDS"))
#excess_birth_year_month_age_reg_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_age_reg_df",".RDS"))

excess_birth_year_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_ctz_df",".RDS"))
excess_birth_year_age_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_age_ctz_df",".RDS"))
excess_birth_year_month_ctz_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_ctz_df",".RDS"))
#excess_birth_year_month_age_ctz_df = readRDS(paste0(code_root_path,results/",save.date,"_",mod_name,"_","seedid",seed_id,"_","excess_birth_year_month_age_ctz_df",".RDS"))

#load results
par_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_par.RDS"))
birth_prob_by_age_df=  readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_birthprob.RDS"))
gp_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_gp.RDS"))
age_bias_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_agebias.RDS"))
pred_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_pred.RDS"))
par_df = readRDS(paste0(code_root_path,"results/",save.date,"_",mod_name,"_seedid",seed_id,"_par.RDS"))
```

```{r}
par_df %>%
  filter(str_detect(variable, "gamma_month")) %>% 
  tidyr::extract(variable,into=c("variable","month_id"),
                 regex =paste0('(\\w.*)\\[',paste(rep("(.*)",1),collapse='\\,'),'\\]'), remove = T) %>%
  as_tibble() %>%
  dplyr::select(month_id,est=mean,lwb=`2.5%`,upb=`97.5%`) %>%
  dplyr::mutate(month_id=as.numeric(month_id)) %>% 
  ggplot(aes(x=month_id,y=est,ymin=lwb,ymax=upb)) +
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")
```

```{r}
birth_prob_by_age_df %>% 
    filter(year %in% c(2000,2010,2024)) %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
    geom_line(aes(col=factor(year)))
```

```{r}
gp_df %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="blue",alpha=0.2)+
  geom_line(col="blue")
```

```{r}
age_bias_df %>% 
  ggplot(aes(x=mother_age))+
  geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
  geom_line(aes(y=est),col="darkred")
    
```

```{r,fig.height=12,fig.width=10}
pred_df %>% 
  filter(age_id %in% (c(15,16,18,19,20,24,26)-14)) %>% 
  ggplot(aes(x=year_id))+
  geom_ribbon(aes(ymin=lwb,ymax=upb),alpha=0.2,fill="darkred")+
  geom_line(aes(y=est),col="darkred")+
  geom_point(aes(y=n_birth),size=2)+
  facet_grid(mother_age~.,scale="free_y")
```

### Excess mortality

#### national, year

```{r}
#national, year
plot_df = excess_birth_year_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~.,scale="free_y")
```

Citizenship

```{r}
#national, year
plot_df = excess_birth_year_ctz_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~citizenship,scale="free_y")
```

#### national, year, month

```{r}
#national, year, month
plot_df = excess_birth_year_month_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year>=2020) %>% 
  ggplot(aes(x = date, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>%  filter(year>=2020) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>%  filter(year>=2020), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~.,scale="free_y")

```

Citizenship

```{r}
#national, year, month
plot_df = excess_birth_year_month_ctz_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year>=2020) %>% 
  ggplot(aes(x = date, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>%  filter(year>=2020) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>%  filter(year>=2020), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~citizenship,scale="free_y")

```

#### national, age at given year

```{r}
#national, age at given year
plot_df = excess_birth_year_age_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year %in% c(2016:2019)) %>% 
  ggplot(aes(x = age, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>% filter(year %in% c(2016:2019)) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>% filter(year %in% c(2016:2019)), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~year,scale="free_y")

```

```{r}
#national, age at given year
plot_df %>% 
  filter(year %in% c(2020:2024)) %>% 
  ggplot(aes(x = age, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>% filter(year %in% c(2020:2024)) , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>% filter(year %in% c(2020:2024)), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~year,scale="free_y")

```

Citizenship

```{r}
#national, age at given year
plot_df = excess_birth_year_age_ctz_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  filter(year %in% c(2016:2024),citizenship=="swiss") %>% 
  ggplot(aes(x = age, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>% filter(year %in% c(2016:2024), citizenship=="swiss"), var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>% filter(year %in% c(2016:2024), citizenship=="swiss"),var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~year,scale="free_y")

```

```{r}
#national, age at given year
plot_df %>% 
  filter(year %in% c(2016:2024),citizenship=="non-swiss") %>% 
  ggplot(aes(x = age, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df %>% filter(year %in% c(2016:2024),citizenship=="non-swiss") , var %in% c("Excess birth", "Relative excess")),
             aes(yintercept = 0),
             linetype = 2) +
  geom_line(data = subset(plot_df  %>% filter(year %in% c(2016:2024),citizenship=="non-swiss"), var %in% c("Expected birth")),
            aes(y=n_birth),col="darkred",
            linetype = 2 ) +
  facet_grid(var~year,scale="free_y")

```

#### By year and district

```{r, fig.height=16,fig.width=12}
#load region location
regions_sf <- st_read(paste0(code_root_path,"data/boundary_data/Boundaries_K4_District_20260101.shp")) #source: https://www.agvchapp.bfs.admin.ch/fr/boundaries?SnapshotDate=01.01.2018&Unit=BAE2018
regions_sf = regions_sf[,c(1,3,7)]
names(regions_sf)[c(1,2)] <- c("dist_name2", "dist_id")

excess_birth_year_reg_df %>% 
  filter(year %in% 2017:2024) %>% 
  dplyr::mutate(rel_exc_mean = n_exc_mean/n_exp_mean) %>% 
  left_join(regions_sf %>% dplyr::mutate(dist_id=as.numeric(dist_id)), by = c("dist_id")) %>%
  st_as_sf() %>%
  ggplot(aes(fill = rel_exc_mean)) +
  geom_sf(color = "black") +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",        # negative
    mid = "white",      # zero
    high = "green",     # positive
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1)
  )+
  theme( legend.position = "bottom",
         legend.direction = "horizontal")+
  facet_wrap(year~.,ncol=2)

```

```{r,include=FALSE}
if(FALSE){
  #other
  excess_birth_year_df %>% 
    ggplot(aes(x = year, y = n_exp_mean, ymin = n_exp_lwb, ymax = n_exp_upb)) +
    geom_ribbon( alpha = 0.2) +geom_line()
  excess_birth_year_df %>% 
    ggplot(aes(x = year, y = n_exp_mean/n_pop, ymin = n_exp_lwb/n_pop, ymax = n_exp_upb/n_pop)) +
    geom_ribbon( alpha = 0.2) +geom_line()
  
  #district
  df = excess_birth_year_reg_df %>% 
    filter(dist_id %in% sample(unique(excess_birth_year_reg_df$dist_id), 6)) %>% 
    left_join(excess_birth_year_df %>% dplyr::select(year,n_pop_nat=n_pop,n_exp_mean_nat=n_exp_mean),by="year")
  df %>% 
    ggplot(aes(x = year, y = n_exp_mean, ymin = n_exp_lwb, ymax = n_exp_upb)) +
    geom_ribbon(aes(fill=dist_name), alpha = 0.2) +
    geom_line(aes(col=dist_name)) +
    facet_grid(dist_name~.,scale="free_y")
  df %>% 
    ggplot(aes(x = year, y = n_exp_mean/n_pop, ymin = n_exp_lwb/n_pop, ymax = n_exp_upb/n_pop)) +
    geom_ribbon(aes(fill=dist_name), alpha = 0.2) +
    geom_line(aes(col=dist_name)) +
    geom_line(aes(y=n_exp_mean_nat/n_pop_nat),lty=2)+
    facet_grid(dist_name~.)
  excess_birth_year_reg_df %>% 
    ggplot(aes(x = year, y = n_exp_mean/n_pop)) +
    geom_line(aes(col=dist_name)) +
    expand_limits(y=0)+theme(legend.position = "none")
}

```

increase subset (now 100) note that district level excess is only available from 2011 as district pop data only for 2011-2025
