---
title: "Birth decline in Switzerland: supplementary figures"
format:
  html:
    theme: cosmo
    code-fold: true        # adds "Show/Hide Code" buttons
    code-summary: "Code"   # customize button text
    toc: true
    toc-location: left  # place TOC on the left
    number-sections: true  # number sections
    toc-depth: 3      # show subsections up to 3 levels
    self-contained: true
    toc-fold: true     # allow collapsing/expanding sections
execute:
  echo: false
  fig-width: 7
  fig-height: 5
  warning: false
  message: true
  comment: ""
params:
  save.date: "20260309"
  save.date2: "20260320"
  seed_id: 1
  res_path: "results/"
  ntile_year_suffix: ""
  last_year: 2024

editor: visual
---


```{r, include=FALSE}
source("../R/000_setup.R")

save.date         = params$save.date
save.date2        = params$save.date2
seed_id           = params$seed_id
res_path          = params$res_path
ntile_year_suffix = params$ntile_year_suffix
last_year         = params$last_year

subgroup_names = c("Main analyis", "Swiss","Non-Swiss","First births","Second and higher-order births")

make_mod_name = function(filter_ctz, filter_parity, last_year){
  mod_name = "mod8"
  mod_name = paste0(mod_name, ifelse(length(filter_ctz)==2, "", paste0("_", filter_ctz)))
  mod_name = if_else(filter_parity=="all", mod_name, paste0(mod_name, "_", filter_parity))
  mod_name = if_else(last_year!=2025, mod_name, paste0(mod_name, "_", last_year))
  mod_name
}

# configs in display order: main, swiss, non-swiss, first, second
mod_df = data.frame(
  ctz_name      = c("",        "swiss", "non-swiss", "",       ""),
  filter_parity = c("all",     "all",   "all",       "first",  "second"),
  stringsAsFactors = FALSE
)
mod_df$mod_name  = mapply(function(ctz, par) make_mod_name(if(ctz=="") c("swiss","non-swiss") else ctz, par, last_year),
                          mod_df$ctz_name, mod_df$filter_parity)
mod_df$save.date = ifelse(mod_df$filter_parity != "all", save.date2, save.date)
mod_df$seed_id   = seed_id

gamma_month_df_list = list()
birth_prob_by_age_df_list = list()
gp_rel_df_list = list()
excess_birth_year_df_list = list()
excess_birth_year_month_df_list = list()
excess_birth_adj2_mun_df_list = list()
excess_birth_year_adj2_mun_df_list = list()
excess_by_ntiles_df_list = list()
for(i in 1:5){
  gamma_month_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_gammamonth.RDS"))
  birth_prob_by_age_df_list[[i]] =  readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_birthprob.RDS"))
  gp_rel_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_gp_rel.RDS"))

  excess_birth_year_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_","excess_birth_year_df",".RDS"))
  excess_birth_year_month_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_","excess_birth_year_month_df",".RDS"))

  excess_birth_adj2_mun_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],ifelse(i %in% c(4,5),"_childless",""),
                                                "_seedid",mod_df$seed_id[i],
                                                "_","excess_birth_adj2_mun_df",".RDS"))
  excess_birth_year_adj2_mun_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],ifelse(i %in% c(4,5),"_childless",""),
                                                     "_seedid",mod_df$seed_id[i],
                                                     "_","excess_birth_year_adj2_mun_df",".RDS"))

  excess_by_ntiles_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],ntile_year_suffix,"_","excess_birth_ntiles_df",".RDS"))
}

#excess by ctz region, only for main analysis and analysis filter on non-swiss
excess_birth_ctzreg_df_list = list()
excess_birth_year_ctzreg_df_list = list()
for(i in c(1,3)){
  excess_birth_ctzreg_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_","excess_birth_ctzreg_df",".RDS"))
  excess_birth_year_ctzreg_df_list[[i]] = readRDS(paste0(code_root_path,res_path,mod_df$save.date[i],"_",mod_df$mod_name[i],"_seedid",mod_df$seed_id[i],"_","excess_birth_year_ctzreg_df",".RDS"))
}


#population (women 15-50) by country
pop_detctz_df = readRDS(paste0(code_root_path,"savepoint/","pop_detctz_country_df.RDS"))

```


```{r, include=FALSE}
fig1_txt = 'p1 = excess_birth_year_month_df %>% 
  ggplot(aes(x=date))+
  geom_ribbon(aes(ymin=n_exp_lwb, ymax=n_exp_upb),alpha=0.2)+
  geom_line(aes(y=n_exp_mean)) +
  geom_point(aes(y=n_birth), alpha=0.6,col="darkred")+
  scale_x_date(name="", expand = c(0.01, 0.01))+
  scale_y_continuous(name="Monthly births")

#fertility
p2 = birth_prob_by_age_df %>% 
    filter(year %in% c(2000,2010,last_year)) %>%
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(year)),alpha=0.2)+
    geom_line(aes(col=factor(year)))+
    scale_x_continuous(name="Maternal age")+
    scale_y_continuous(name="Probability of having a child\n(by month)",labels = scales::percent)+
    scale_color_discrete(name="Calendar year")+
    scale_fill_discrete(name="Calendar year")


#calculate peak age
ref_year = min(birth_prob_by_age_df$year)
df <- birth_prob_by_age_df %>% filter(year==ref_year) %>%  arrange(mother_age)
idx <- which.max(df$est)
y1 <- df$est[idx - 1]
y2 <- df$est[idx]
y3 <- df$est[idx + 1]
x0 <- df$mother_age[idx]
delta <- (y1 - y3) / (2 * (y1 - 2 * y2 + y3))
peak_age <- x0 + delta

p3 = gp_rel_df %>% 
  dplyr::mutate(est = est + peak_age,
                lwb = lwb + peak_age,
                upb = upb + peak_age) %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(fill="black",alpha=0.2)+
  geom_line(col="black")+
  scale_y_continuous(name="Age of fertility peak")+
  scale_x_continuous(name="Calendar year")

cowplot::plot_grid(p1,
                   cowplot::plot_grid(p2,p3,rel_widths = c(1.4,1),ncol=2,labels = c("B.","C.")),
                   rel_heights = c(1,1),labels = c("A.",""),
                   nrow = 2)'


fig1b_txt = 'gamma_month_df = gamma_month_df %>% 
  dplyr::mutate(month_abb = factor(month_id, levels = 1:12, labels = month.abb)) 
  gamma_month_df %>% 
  ggplot(aes(x = month_id,
  y = est,ymin = lwb,ymax = upb)) +
  geom_ribbon(fill = "blue", alpha = 0.2) +
  geom_line(col = "blue", group = 1) +
  scale_x_continuous(name="Month",breaks=1:12,labels=gamma_month_df$month_abb)+
  scale_y_continuous(name = "Relative effect on birth (vs January)",labels=scales::percent)'

fig2_txt = '
#national, year
plot_df1 = excess_birth_year_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))

p1 = plot_df1 %>% 
  filter(var %in% c("Expected birth")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean)) +
  geom_point(aes(y=n_birth),col="darkred" ) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Annual birth")

p2 = plot_df1 %>% 
  filter(var %in% c("Relative excess")) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean))+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_continuous(name="Year")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)

#national, year, month
plot_df2 = excess_birth_year_month_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
p3 = plot_df2 %>% 
  filter(year>=2020, var=="Relative excess") %>% 
    ggplot(aes(x = date)) +
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line(aes(y = mean))+
  geom_hline( aes(yintercept = 0),    linetype = 2) +
  scale_x_date(name="")+
  scale_y_continuous(name="Relative excess birth",labels=scales::percent)
cowplot::plot_grid(p1,
                                      p2,
                                      p3,
                                      nrow=3,labels=c("A.","B.","C."))'

fig3_txt = 'excess_birth_adj2_mun_df %>% 
  #left_join(new_mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  left_join(new_mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>% 
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean), color = NA) +  # communes sans bordure
  geom_sf(data = regions_sf %>% mutate(dist_id = as.numeric(dist_id)), 
          fill = NA, color = "black", size = 0.3) +  # contours districts
  geom_sf(data = lake_sf, fill = "lightblue", color = NA, alpha = 0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",
    mid = "lightyellow",
    high = "green",
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1),
    #limits=c(-0.4,0.4)
  ) +
  theme(legend.position = "bottom",
        legend.direction = "horizontal")'

fig4_txt ='
childless_plot = ifelse(length(unique(excess_by_ntiles_df$childless))==2,TRUE,FALSE)
excess_by_ntiles_df %>% 
  left_join(explanatory_var_df %>% 
              dplyr::mutate(explanatory_var = paste0(explanatory_var,"_ntile")),by="explanatory_var") %>% 
  filter(!is.na(explanatory_var2), childless==childless_plot) %>% 
  dplyr::mutate(explanatory_var2 = factor(explanatory_var2, levels = explanatory_var_df$explanatory_var2)) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(alpha = 0.2) +
  geom_line() +
  geom_point() +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_x_continuous(name="Population decile",breaks=c(1:10))+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var2~.,ncol=2)'

fig5_txt = '
rel_pop_detctz_df = pop_detctz_df %>% 
  left_join(ctz_map,by="ctz_name") %>% 
  group_by(ctz_region,year) %>% 
  dplyr::summarise(n=sum(n),.groups="drop_last") %>% 
  dplyr::summarise(n=mean(n),.groups="drop") %>% 
  dplyr::mutate(p=n/sum(n))
region_levels <- excess_birth_year_ctzreg_df %>%
  left_join(rel_pop_detctz_df, by="ctz_region") %>%
  distinct(ctz_region, p) %>%
  arrange(-p) %>%
  mutate(ctz_region2 = paste0(ctz_region, " (", round(p*100, 1), "%)")) %>%
  pull(ctz_region2)

n=15
cols <- scales::hue_pal()(n)
eu_colors <- cols[c(1,3,5)] #scales::seq_gradient_pal("darkblue", "#ABD9E9")(seq(0,1,length.out=3))
rest_colors <- cols[c(7,8,9,10,14,15)]# scales::seq_gradient_pal("#D7191C", "#FDAE61")(seq(0,1,length.out=6))

world_reg_name = c("Switzerland", "Western Europe", "Eastern Europe",
                   "South America", "North America", "Central America",
                   "Africa", "Asia", "Oceania")

rel_pop_detctz_df <- rel_pop_detctz_df %>%
  mutate(ctz_region2 = paste0(ctz_region, " (", round(p*100, 1), "%)"),
         ctz_region2 = factor(ctz_region2, levels = region_levels)) %>%
  #sort so that colors are in right orders
  mutate(ctz_region = factor(ctz_region, levels = world_reg_name),
         ctz_region3 = gsub(" ", "\n", ctz_region),
         ctz_region4 = gsub(" ", "\n", ctz_region2)) %>% 
  arrange(ctz_region)
rel_pop_detctz_df$fill_col = c(eu_colors,rest_colors)

p4 = excess_birth_ctzreg_df %>%
  left_join(rel_pop_detctz_df,by="ctz_region") %>% 
  ggplot(aes(x=ctz_region4,y=rel_exc_mean,ymin=rel_exc_lwb,ymax=rel_exc_upb,col=ctz_region4))+
  geom_hline(yintercept = 0, lty=2)+
  geom_pointrange()+
  scale_x_discrete(name="",
                   limits = rel_pop_detctz_df$ctz_region4)+
  scale_y_continuous(name = "Relative excess birth",breaks=c(-0.2,0,0.2,0.4,0.6,0.8,1),labels = scales::percent)+
  scale_color_manual( name = "Region",
                      values = rel_pop_detctz_df$fill_col,
                      breaks = rel_pop_detctz_df$ctz_region4) +
  theme(legend.position = "none")
p5 = excess_birth_year_ctzreg_df %>%
  left_join(rel_pop_detctz_df, by="ctz_region") %>% 
  ggplot(aes(x=year, y=rel_exc_mean, ymin=rel_exc_lwb, ymax=rel_exc_upb)) +
  geom_ribbon(aes(fill=ctz_region), alpha=0.1) +
  geom_line(aes(col=ctz_region)) +
  facet_wrap(ctz_region~.,nrow=3) +
  geom_hline(yintercept = 0, lty=2) +
  scale_y_continuous(name = "Relative excess birth",breaks=c(-0.2,0,0.2,0.4,0.6,0.8,1),labels = scales::percent)+
  scale_x_continuous(name="",breaks=c(2011,2015,2020,last_year))+
  scale_color_manual( name = "",
                      values = rel_pop_detctz_df$fill_col,
                      breaks = rel_pop_detctz_df$ctz_region)+
  scale_fill_manual( name = "",
                     values = rel_pop_detctz_df$fill_col,
                     breaks = rel_pop_detctz_df$ctz_region)+
  theme(legend.position = "bottom",
        legend.box.margin = margin(t = -18))

cowplot::plot_grid(p4,
                   p5+theme(legend.position = "none"),
                   get_legend2(p5),
                   nrow=3,labels=c("A.","B.",""),rel_heights = c(1,2,0.2))'

```

```{r, results='asis'}
library(knitr)

for (i in 1:5) {
  gamma_month_df= gamma_month_df_list[[i]]
  excess_birth_year_df = excess_birth_year_df_list[[i]]
  excess_birth_year_month_df = excess_birth_year_month_df_list[[i]]
  birth_prob_by_age_df = birth_prob_by_age_df_list[[i]]
  gp_rel_df = gp_rel_df_list[[i]]
  excess_birth_adj2_mun_df = excess_birth_adj2_mun_df_list[[i]]
  excess_by_ntiles_df = excess_by_ntiles_df_list[[i]]
  if(i %in% c(1,3)){
    excess_birth_ctzreg_df = excess_birth_ctzreg_df_list[[i]]
    excess_birth_year_ctzreg_df = excess_birth_year_ctzreg_df_list[[i]]
  }

  cat(paste0("# ", subgroup_names[i]))
  
  if(i>1){
  txt <- c(
    paste0("## Model fit"),
    "",
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot1_", i),
    
    paste0('#| fig-cap: "**Model fit**. A. Monthly number of births, as observed (red points) and expected (black lines corresponding to the mean posterior and gray area to the 95% credibility interval (95% CrI)), from 2000 to ', last_year, '. B. Fertility curve. Probability of having a child (by month) according to maternal age for different calendar years (2000, 2010 and ', last_year, '). C. Age of fertility peak according to calendar year."'),
    "#| fig-width: 10",
    "#| fig-height: 6",
    "",
    
    fig1_txt,
    
    "```",
    "")

  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  }
   txt <- c(
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot1b_", i),
    
    '#| fig-cap: "**Model fit**. Effect of month on the probability of having a child."',
    "#| fig-width: 5",
    "#| fig-height: 4",
    "",
    
    fig1b_txt,
    
    "```",
    "")

  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  
  if(i>1){
  txt <- c(
   paste0("## Excess birth"),
   "",
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot2_", i),

    paste0('#| fig-cap: "**Excess birth**. A. Annual number of births, as observed (red points) and expected (black lines corresponding to the mean posterior and gray area to the 95% credibility interval (95% CrI)), from 2000 to ', last_year, '. B. Relative excess birth by year from 2000 to ', last_year, '. C. Relative excess birth by month between 2020 and ', last_year, '."'),
    "#| fig-width: 10",
    "#| fig-height: 6",
    "",

    fig2_txt,

    "```",
    "")

  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  
  txt <- c(
   paste0("## Excess birth by municipality"),
   "",
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot3_", i),

    paste0('#| fig-cap: "**Excess birth by municipality**. Relative excess birth by municipality between 2017 and ', last_year, '."'),
    "#| fig-width: 10",
    "#| fig-height: 6",
    "",

    fig3_txt,

    "```",
    "")

  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  }
  
   txt <- c(
   paste0("## Excess birth by variable deciles"),
   "",
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot4_", i),

    '#| fig-cap: "**Relative excess birth stratified by variable deciles**"',
    "#| fig-width: 10",
    "#| fig-height: 14",
    "",

    fig4_txt,

    "```",
    "")
   
  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  
  if(i %in% c(1,3)){
    txt <- c(
   paste0("## Excess birth by maternal citizenship"),
   "",
   "```{r, echo=FALSE}",
    paste0("#| label: fig-plot5_", i),

    paste0('#| fig-cap: "**Excess birth by maternal citizenship**: A. Relative excess birth by citizenship aggregated by region of the world in 2011-', last_year, '. B. Relative excess birth by citizenship over years."'),
    "#| fig-width: 8",
    "#| fig-height: 10",
    "",

    fig5_txt,

    "```",
    "")
    
  cat("\n\n")
  cat(knit_child(text = txt, quiet = TRUE))
  cat("\n\n")
  }
  
}
```






