### Fertility curve

```{r}
rbind(birth_prob_by_age_df2 %>% filter(year==2000) %>% dplyr::mutate(ctz="swiss"),
      birth_prob_by_age_df3 %>% filter(year==2000) %>% dplyr::mutate(ctz="non-swiss")) %>% 
    ggplot(aes(x=mother_age,y=est,ymin=lwb,ymax=upb))+
    geom_ribbon(aes(fill=factor(ctz)),alpha=0.2)+
    geom_line(aes(col=factor(ctz)))+
    scale_y_continuous(labels = scales::percent)
```


### Age shift

```{r}
rbind(gp_rel_df2 %>% dplyr::mutate(ctz="swiss"),
      gp_rel_df3 %>% dplyr::mutate(ctz="non-swiss")) %>% 
  ggplot(aes(x=year,y=est,ymin=lwb,ymax=upb))+
  geom_ribbon(aes(fill=ctz),alpha=0.2)+
  geom_line(aes(col=ctz))+
  scale_y_continuous(name="Change in age of fertility peak (relative to 2000)")
```

```{r}
rbind(gamma_month_df2 %>% dplyr::mutate(ctz="swiss"),
      gamma_month_df3 %>% dplyr::mutate(ctz="non-swiss")) %>% 
  ggplot(aes(x=month_id,y=est,ymin=lwb,ymax=upb,fill=ctz)) +
  geom_ribbon(alpha=0.2)+
  geom_line(aes(col=ctz))
```

#### Swiss

```{r}
#national, year
plot_df = excess_birth_year_df2 %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~.,scale="free_y")
```

#### Non-wiss

```{r}
#national, year
plot_df = excess_birth_year_df3 %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
             n_exc_mean, n_exc_lwb, n_exc_upb,
             rel_exc_mean, rel_exc_lwb, rel_exc_upb),
    names_to = c("var", "stat"),
    names_pattern = "(.*)_(mean|lwb|upb)",
    values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
plot_df %>% 
  ggplot(aes(x = year, y = mean))+
  geom_ribbon(aes(ymin = lwb,ymax=upb),alpha = 0.2) +
  geom_line() +
  geom_hline(data = subset(plot_df, var %in% c("Excess birth", "Relative excess")),
    aes(yintercept = 0),
    linetype = 2) +
  geom_line(data = subset(plot_df, var %in% c("Expected birth")),
    aes(y=n_birth),col="darkred",
    linetype = 2 ) +
  facet_grid(var~.,scale="free_y")
```



#### Swiss

```{r, fig.height=6,fig.width=8}
excess_birth_adj2_mun_df2 %>% 
  left_join(new_mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  left_join(new_mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>% 
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean), color = NA) +  # communes sans bordure
  geom_sf(data = regions_sf %>% mutate(dist_id = as.numeric(dist_id)), 
          fill = NA, color = "black", size = 0.3) +  # contours districts
  geom_sf(data = lake_sf, fill = "lightblue", color = NA, alpha = 0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",
    mid = "lightyellow",
    high = "green",
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1),
    #limits=c(-0.4,0.4)
  ) +
  theme(legend.position = "bottom",
        legend.direction = "horizontal")
```

**Figure XX**: Smoothed estimates of excess birth by municipality for 2017-2024.

#### Non-Swiss

```{r, fig.height=6,fig.width=8}
excess_birth_adj2_mun_df3 %>% 
  left_join(new_mun_df %>% dplyr::select(mun_id,dist_id) %>% distinct(), by="mun_id") %>% 
  left_join(new_mun_sf %>% dplyr::mutate(mun_id=as.numeric(mun_id)), by = c("mun_id")) %>% 
  st_as_sf() %>%
  ggplot() +
  geom_sf(aes(fill = rel_exc_mean), color = NA) +  # communes sans bordure
  geom_sf(data = regions_sf %>% mutate(dist_id = as.numeric(dist_id)), 
          fill = NA, color = "black", size = 0.3) +  # contours districts
  geom_sf(data = lake_sf, fill = "lightblue", color = NA, alpha = 0.5) +
  scale_fill_gradient2(
    name = "Relative excess birth",
    low = "red",
    mid = "lightyellow",
    high = "green",
    midpoint = 0,
    labels = scales::percent_format(accuracy = 1),
    #limits=c(-0.4,0.4)
  ) +
  theme(legend.position = "bottom",
        legend.direction = "horizontal")
```

**Figure XX**: Smoothed estimates of excess birth by municipality for 2017-2024.






### Swiss

```{r,fig.width=8,fig.height=6}
df_var = excess_birth_year_mun_draw_df2 %>% dplyr::select(mun_id,mun_name) %>% distinct() %>% 
  left_join(new_pop_mun_df %>% filter(year==2024,month==1) %>% 
              group_by(mun_id) %>% dplyr::summarise(n_pop = sum(n),.groups="drop"), by=c("mun_id")) %>% 
  left_join(rural_urban_df %>% dplyr::select(mun_id,urban_rural_type2),by="mun_id") %>% 
  left_join(pop_dens_df %>% pivot_wider(id_cols = mun_id,names_from = "variable",values_from = "value"),by="mun_id") %>% 
  left_join(sep_df3 %>% dplyr::select(mun_id,sep=ssep3_d_mean),by="mun_id")


make_pop_ntile <- function(data, var, n = 5) {
  total_pop = sum(data$n_pop)
  data %>%
    arrange({{ var }}) %>%
    dplyr::mutate(cum_pop = cumsum(n_pop),
                  group = cut(cum_pop,breaks = seq(0, total_pop, length.out = n+1),labels = FALSE,  include.lowest = TRUE )) %>% 
    dplyr::select(-cum_pop) %>% 
    arrange(mun_id) %>%
    pull(group)
}

vars <- setdiff(names(df_var), c("urban_rural_type2", "mun_id", "mun_name"))

for (v in vars) {
  df_var[[paste0(v, "_ntile")]] <-make_pop_ntile(df_var, !!rlang::sym(v), 10)
}
df_var$urban_rural_type2_ntile = dense_rank(df_var$urban_rural_type2)

if(FALSE){
  #check that the number of pop is more or less the same between groups (except for urban_rural_type2)
  df_var %>% group_by(n_pop_ntile) %>% 
    dplyr::summarise(n_pop = sum(n_pop),
                     n=n())
}

#aggregate excess by group
excess_df = list()
for(v in paste0(c("urban_rural_type2",vars), "_ntile")){
  excess_df[[v]] = get_excess_est(var_group=c("ntile"),
                                  excess_birth_year_mun_draw_df2 %>% 
                                    filter(year %in% 2017:2024) %>% 
                                    left_join(df_var %>% dplyr::select(mun_id,ntile = !!v),by="mun_id")) %>% 
    dplyr::mutate(explanatory_var=!!v)
}
excess_df = rbindlist(excess_df)


#plot
plot_df = excess_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
excess_df %>% 
  filter(explanatory_var %in% c("n_pop_ntile","pop_dens_building_ntile", "prop_building_ntile", "prop_individual_houses_ntile","sep_ntile","urban_rural_type2_ntile")) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(aes(),alpha = 0.2) +
  geom_line() +
  geom_point() +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var~.,ncol=2)
```

**Figure XX**: Relative excess birth estimates by decile of explanatory variables.

### Non-Swiss

```{r,fig.width=8,fig.height=6}
df_var = excess_birth_year_mun_draw_df3 %>% dplyr::select(mun_id,mun_name) %>% distinct() %>% 
  left_join(new_pop_mun_df %>% filter(year==2024,month==1) %>% 
              group_by(mun_id) %>% dplyr::summarise(n_pop = sum(n),.groups="drop"), by=c("mun_id")) %>% 
  left_join(rural_urban_df %>% dplyr::select(mun_id,urban_rural_type2),by="mun_id") %>% 
  left_join(pop_dens_df %>% pivot_wider(id_cols = mun_id,names_from = "variable",values_from = "value"),by="mun_id") %>% 
  left_join(sep_df3 %>% dplyr::select(mun_id,sep=ssep3_d_mean),by="mun_id")


make_pop_ntile <- function(data, var, n = 5) {
  total_pop = sum(data$n_pop)
  data %>%
    arrange({{ var }}) %>%
    dplyr::mutate(cum_pop = cumsum(n_pop),
                  group = cut(cum_pop,breaks = seq(0, total_pop, length.out = n+1),labels = FALSE,  include.lowest = TRUE )) %>% 
    dplyr::select(-cum_pop) %>% 
    arrange(mun_id) %>%
    pull(group)
}

vars <- setdiff(names(df_var), c("urban_rural_type2", "mun_id", "mun_name"))

for (v in vars) {
  df_var[[paste0(v, "_ntile")]] <-make_pop_ntile(df_var, !!rlang::sym(v), 10)
}
df_var$urban_rural_type2_ntile = dense_rank(df_var$urban_rural_type2)

if(FALSE){
  #check that the number of pop is more or less the same between groups (except for urban_rural_type2)
  df_var %>% group_by(n_pop_ntile) %>% 
    dplyr::summarise(n_pop = sum(n_pop),
                     n=n())
}

#aggregate excess by group
excess_df = list()
for(v in paste0(c("urban_rural_type2",vars), "_ntile")){
  excess_df[[v]] = get_excess_est(var_group=c("ntile"),
                                  excess_birth_year_mun_draw_df3 %>% 
                                    filter(year %in% 2017:2024) %>% 
                                    left_join(df_var %>% dplyr::select(mun_id,ntile = !!v),by="mun_id")) %>% 
    dplyr::mutate(explanatory_var=!!v)
}
excess_df = rbindlist(excess_df)


#plot
plot_df = excess_df %>% 
  pivot_longer(cols = c(n_exp_mean, n_exp_lwb, n_exp_upb,
                        n_exc_mean, n_exc_lwb, n_exc_upb,
                        rel_exc_mean, rel_exc_lwb, rel_exc_upb),
               names_to = c("var", "stat"),
               names_pattern = "(.*)_(mean|lwb|upb)",
               values_to = "value" ) %>% 
  pivot_wider(names_from = "stat",values_from = "value") %>% 
  dplyr::mutate(var = factor(var,levels=c("n_exp","n_exc","rel_exc"),labels=c("Expected birth","Excess birth","Relative excess")))
excess_df %>% 
  filter(explanatory_var %in% c("n_pop_ntile","pop_dens_building_ntile", "prop_building_ntile", "prop_individual_houses_ntile","sep_ntile","urban_rural_type2_ntile")) %>% 
  ggplot(aes(x = ntile, y = rel_exc_mean, ymin = rel_exc_lwb, ymax = rel_exc_upb))+
  geom_ribbon(aes(),alpha = 0.2) +
  geom_line() +
  geom_point() +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  scale_y_continuous(name="Relative excess birth",labels = scales::percent)+
  facet_wrap(explanatory_var~.,ncol=2)
```

**Figure XX**: Relative excess birth estimates by decile of explanatory variables.
